
% qa_certificate_paper_skeleton.tex
% Skeleton: Failure-Aware Reachability Certificates for QA-AlphaGeometry + Physics Projection

\documentclass[11pt]{article}
\usepackage{amsmath,amssymb}
\usepackage{hyperref}
\usepackage{geometry}
\geometry{margin=1in}

\title{Failure-Aware Reachability Certificates for QA: \\ Unified Proof and Projection Artifacts}
\author{Will Dale}
\date{January 2026}

\begin{document}
\maketitle

\begin{abstract}
We introduce a \emph{proof certificate} schema for Quantum Arithmetic (QA) that treats both
successful derivations and certified obstructions as first-class mathematical objects.
Unlike standard proof traces which only show success, our certificates always emit a
reviewable obstruction witness when search fails (e.g., depth exhaustion, invariant violation,
component disconnection).
We further extend the same certificate machinery to projection-dependent physics
experiments, where classical laws emerge only under specific observers.
All claims are packaged as deterministic JSON artifacts.
\end{abstract}

\section{Motivation}
Two failures repeatedly block reproducible claims in symbolic / geometric proof search:
(i) successes are traceable, but failures are not, and (ii) ``physics-like'' behavior in QA
is often asserted without a falsifiable projection contract.
We address both by standardizing \emph{certificates}.

\section{Certificate Objects}
Define a certificate as a tuple:
\[
\mathcal{C} = (\texttt{generator\_set}, \texttt{contracts}, \texttt{witness\_type}, \texttt{witness})
\]
where all reachability statements are relative to a generator universe.

\subsection{Generators and Namespaces}
QA substrate generators use the canonical set $\{\sigma,\lambda,\mu,\nu\}$ (and optional inverses).
External engines are namespaced:
\[
\texttt{AG:rule\_id},\quad \texttt{PHYS:probe}.
\]

\subsection{Invariants and Non-Reduction}
Certificates optionally track a list of invariants that must be preserved.
QA's Non-Reduction Axiom is represented as a contract flag; violations are certified
as \texttt{NON\_REDUCTION\_VIOLATION}.

\subsection{Failure Taxonomy}
We treat failure modes as a finite algebra of obstructions (not bugs), including:
\texttt{DEPTH\_EXHAUSTED}, \texttt{GENERATOR\_INSUFFICIENT}, \texttt{SCC\_UNREACHABLE}, etc.

\section{Results: Integration with AlphaGeometry}
\label{sec:alphageometry-results}

We demonstrate the certificate framework by integrating it with
\emph{AlphaGeometry}, a state-of-the-art automated geometry prover.
Our goal is not to modify AlphaGeometry's proof search, but to
\emph{externalize its outcomes}—both successes and failures—into a
uniform, first-class certificate object.

\subsection{Adapter Overview}

AlphaGeometry produces a \texttt{SearchResult} object containing:
(i) a Boolean solved flag, (ii) an optional proof trace, and
(iii) search statistics (depth, expansions, policy).
We implement a deterministic adapter that maps each \texttt{SearchResult}
into a \texttt{ProofCertificate} without altering search behavior.

Key design choices are:

\begin{itemize}
  \item \textbf{Namespaced generators.}
  Each AlphaGeometry rule is represented as a generator
  \texttt{AG:<rule\_id>}, ensuring disjointness from QA-native
  generators (\(\sigma,\lambda,\mu,\nu\)) and physics observers.
  \item \textbf{Exact, minimal state references.}
  AlphaGeometry states lack numeric QA coordinates; we therefore use
  stable hash-based identifiers to reference states without inventing
  geometry.
  \item \textbf{Conservative failure classification.}
  When a search terminates without proof, the adapter produces an
  obstruction certificate but does not claim strong properties
  (e.g., unreachable SCCs) unless explicitly justified by evidence.
\end{itemize}

The adapter produces either a \emph{success certificate} with a witness
path, or an \emph{obstruction certificate} with explicit failure
evidence, using the same schema in both cases.

\subsection{Success Certificate: Parallel Transitivity}

Our first example is a successful AlphaGeometry proof of the
\emph{parallel transitivity} theorem.
The exported certificate
(\texttt{parallel\_transitivity\_proof.cert.json})
contains:

\begin{itemize}
  \item A single-step witness path using generator
  \texttt{AG:parallel\_transitivity}.
  \item Deterministic source and destination state identifiers.
  \item An invariant contract with
  \texttt{non\_reduction\_enforced = false},
  reflecting AlphaGeometry's independent algebraic semantics.
\end{itemize}

Formally, this certificate is a reachability witness:
\[
  s_0 \xrightarrow{\texttt{AG:parallel\_transitivity}} s_1,
\]
where the generator set explicitly records the rule required for the
proof. No additional invariants or observers are assumed.

This artifact demonstrates that trivial and non-trivial AlphaGeometry
proofs can be embedded into the certificate framework without loss of
semantic fidelity.

\subsection{Obstruction Certificate: Unsolvable Configuration}

To demonstrate failure handling, we consider an intentionally unsolvable
geometry configuration.
AlphaGeometry terminates immediately, generating no successors and
producing no proof trace.

The corresponding certificate
(\texttt{unsolvable\_obstruction.cert.json}) is an
\emph{obstruction certificate} with:

\begin{itemize}
  \item \texttt{witness\_type = obstruction},
  \item \texttt{fail\_type = depth\_exhausted},
  \item Explicit evidence: zero successors generated at depth zero.
\end{itemize}

Crucially, the certificate does \emph{not} claim that the target state is
globally unreachable under all generators; it certifies only what is
known:
that under the given generator set and search policy, no progress was
possible.

This conservative classification avoids overclaiming while still
producing a reusable, inspectable artifact.

\subsection{Unified Interpretation}

These two results illustrate the central claim of the framework:

\begin{quote}
\emph{Success and failure are objects of the same type, differing only
in witness content.}
\end{quote}

Both certificates:
(i) use the same schema,
(ii) explicitly record their generator assumptions, and
(iii) are reproducible and machine-checkable.

In §\ref{sec:physics-projection}, we show that the same certificate
structure extends beyond theorem proving, capturing physical laws and
their failures as projection-dependent reachability phenomena.

\section{Physics as Projection}
\label{sec:physics-projection}

We add an explicit \texttt{ProjectionContract} recording:
observer id, time projection, topology/symmetry preservation, and projected observables.
A classical law is then stated as: \emph{the law holds under observer $O$}.

\paragraph{Artifacts.}
\begin{itemize}
  \item \texttt{reflection\_GeometryAngleObserver.success.cert.json}
  \item \texttt{reflection\_NullObserver.obstruction.cert.json}
\end{itemize}

\section{Discussion}
Certificates shift evaluation from narrative claims to artifact inspection:
success witnesses are checkable, and failures become reviewable mathematical statements.
The same mechanism unifies formal proof, reachability, and projection-dependent physical laws.

\section{Conclusion}
QA can support a certificate-first scientific workflow:
every claim ships as deterministic JSON, with a canonical schema and adapters
for multiple domains.

\bibliographystyle{plain}
\bibliography{refs}

\end{document}
