\documentclass[11pt,a4paper]{article}
\usepackage[utf8]{inputenc}
\usepackage{amsmath,amssymb,amsthm}
\usepackage{booktabs}
\usepackage{graphicx}
\usepackage{hyperref}
\usepackage[margin=1in]{geometry}

\newtheorem{theorem}{Theorem}
\newtheorem{definition}{Definition}
\newtheorem{proposition}{Proposition}

\title{Transition-Coherence as a Structural Indicator:\\
  A Domain-General Framework from Quantum Arithmetic}

\author{Will Dale}

\date{\today}

\begin{document}
\maketitle

\begin{abstract}
We introduce the \emph{QA Coherence Index} (QCI), a structural indicator
derived from the transition dynamics of Quantum Arithmetic (QA), and show
it detects system instability across unrelated domains. The QCI measures
the rolling prediction accuracy of the QA Fibonacci-shift operator applied
to topographically clustered multi-channel signals. We establish three
claims. \textbf{(A)}~A local observer (multi-asset returns) produces QCI
that anticorrelates with future volatility (partial $r = -0.22$ beyond
lagged realized volatility, 84\% robustness across 80 configurations,
permutation-validated). \textbf{(B)}~A separate global observer
(cross-asset correlation, PCA concentration) produces QCI with the
\emph{opposite sign} ($r = +0.25$): during stress, coupling structure
tightens. \textbf{(C)}~The gap $\mathrm{QCI}_\mathrm{local} -
\mathrm{QCI}_\mathrm{global}$ achieves partial $r = -0.42$ beyond lagged
volatility (100\% robustness, permutation-validated), capturing stress as
\emph{structural contraction with path disorder}. Cross-domain validation
in EEG seizure classification ($\Delta R^2 = +0.21$, 10 patients), audio
signal analysis (partial $r = +0.75$), and seismology ($r = +0.23$ with
future activity, $\chi^2 = 61$ for regime discrimination) confirms the
architecture is domain-general across four unrelated scientific domains. The core finding: deviation from QA-predicted transition
dynamics carries forward-looking structural information that classical
volatility measures do not fully capture.
\end{abstract}

% ═══════════════════════════════════════════════════════════════════════════
\section{Introduction}
% ═══════════════════════════════════════════════════════════════════════════

The problem of detecting structural regime changes in complex systems ---
financial markets, neural dynamics, acoustic signals --- is central to
multiple scientific disciplines. Classical approaches rely on domain-specific
features: realized volatility in finance, spectral power ratios in
neuroscience, autocorrelation in signal processing. While effective within
their domains, these features lack a unifying mathematical framework that
explains \emph{why} they work and \emph{what structural property} they
measure.

Quantum Arithmetic (QA) is a modular dynamical system operating on integer
state pairs $(b,e) \in \{1,\ldots,m\}^2$ via the Fibonacci-shift operator
$T(b,e) = (e, ((b+e-1) \bmod m) + 1)$. The state space partitions into
three orbit types: \emph{singularity} (period 1, fixed point),
\emph{satellite} (period 8 for $m=24$), and \emph{cosmos} (period 24).
This partition is a discrete analogue of the triune vibratory structure
described by Keely and formalized in Sympathetic Vibratory Physics.

We propose that QA provides the missing unifying framework: continuous
observables enter at the \emph{observer boundary}, are projected into discrete
QA states, and the coherence of subsequent state transitions --- measured by
the T-operator's prediction accuracy --- constitutes a domain-general
structural indicator.

\subsection{Contributions}

The paper makes three empirical claims supported by a unifying principle:

\begin{enumerate}
  \item[\textbf{A.}] \textbf{Local QCI detects path-law breakdown.}
    A local topographic observer (multi-channel returns) produces QCI via
    T-operator prediction accuracy. Low local QCI anticorrelates with future
    volatility (partial $r = -0.22$ beyond lagged RV, 84\% robustness,
    permutation-validated).
  \item[\textbf{B.}] \textbf{Global QCI detects manifold contraction.}
    A separate global observer (correlation, PCA concentration, dispersion)
    produces QCI with the \emph{opposite sign}: high global QCI correlates
    with future volatility ($r = +0.25$). During stress, the coupling
    manifold \emph{contracts} (everything correlates).
  \item[\textbf{C.}] \textbf{QCI gap detects stress as structural divergence.}
    The discrepancy $\text{QCI}_{\text{gap}} = \text{QCI}_{\text{local}} -
    \text{QCI}_{\text{global}}$ achieves partial $r = -0.42$ beyond lagged RV
    at best configuration, with 100\% robustness (16/16 grid) and
    permutation-validated ($p < 0.001$). Stress is the regime where the
    market becomes globally more constrained while locally less lawful.
\end{enumerate}

Cross-domain validation in EEG ($\Delta R^2 = +0.21$, 10 patients) and
audio (partial $r = +0.75$) demonstrates the local observer architecture
is not domain-specific.

% ═══════════════════════════════════════════════════════════════════════════
\section{Background: Quantum Arithmetic}
\label{sec:qa}
% ═══════════════════════════════════════════════════════════════════════════

\begin{definition}[QA State Space]
For modulus $m$, the QA state space is $S_m = \{1,\ldots,m\}^2$.
The Fibonacci-shift operator $T: S_m \to S_m$ is defined by
\[
  T(b,e) = \bigl(e,\; ((b+e-1) \bmod m) + 1\bigr).
\]
\end{definition}

\begin{definition}[QA Norm]
The QA norm $f: S_m \to \mathbb{Z}/m\mathbb{Z}$ is
$f(b,e) = b^2 + be - e^2 \pmod{m}$.
This is the norm form of the golden ratio ring $\mathbb{Z}[\varphi]$.
\end{definition}

\begin{proposition}[Orbit Partition]
For $m = 9$, $S_9$ partitions into exactly three orbit families:
\begin{itemize}
  \item \textbf{Singularity}: 1 state $(9,9)$, period 1
  \item \textbf{Satellite}: 8 states, period 8, symmetric 3D structure
  \item \textbf{Cosmos}: 72 states, period 24, linear 1D structure
\end{itemize}
Total: $1 + 8 + 72 = 81 = 9^2$. The orbit period 24 equals the
Pisano period $\pi(9)$.
\end{proposition}

% ═══════════════════════════════════════════════════════════════════════════
\section{The Topographic Observer Architecture}
\label{sec:observer}
% ═══════════════════════════════════════════════════════════════════════════

\subsection{Observer-Projection Firewall (Theorem NT)}

Continuous measurements enter QA \emph{exactly once} at the observer boundary
and exit \emph{exactly once} at the projection boundary. No continuous
quantity participates in QA dynamics. This firewall is both a correctness
constraint and a design principle.

\subsection{Pipeline}

\begin{enumerate}
  \item \textbf{Observer layer} (continuous permitted):
    Multi-channel signal $\mathbf{x}_t \in \mathbb{R}^d$ is clustered into
    $K$ discrete microstates via $k$-means: $c_t = \text{cluster}(\mathbf{x}_t)$.
  \item \textbf{QA layer} (integer only):
    Microstate transitions map to QA states via a designed mapping
    $\phi: \{0,\ldots,K{-}1\}^2 \to S_m$:
    $(b_t, e_t) = \phi(c_t, c_{t+1})$.
  \item \textbf{T-operator prediction}:
    $(\hat{b}_{t+1}, \hat{e}_{t+1}) = T(b_t, e_t)$.
  \item \textbf{Projection layer} (continuous permitted):
    Rolling prediction accuracy over window $W$:
    \[
      \text{QCI}_t = \frac{1}{W} \sum_{s=t-W+1}^{t}
        \mathbf{1}[\hat{e}_{s+1} = e_{s+1}^{\text{actual}}]
    \]
\end{enumerate}

\subsection{Mapping Design}

The microstate-to-QA mapping $\phi$ must ensure all three orbit types
(singularity, satellite, cosmos) are reachable from the microstate transition
set. For $m = 24$, satellite requires multiples of 8; singularity requires
$(24,24)$. Our canonical mapping for $K=6$:
$\{0 \mapsto 8,\; 1 \mapsto 16,\; 2 \mapsto 24,\; 3 \mapsto 5,\;
4 \mapsto 3,\; 5 \mapsto 11\}$.

% ═══════════════════════════════════════════════════════════════════════════
\section{Finance: QCI Predicts Future Volatility}
\label{sec:finance}
% ═══════════════════════════════════════════════════════════════════════════

\subsection{Data and Setup}

Six assets (SPY, QQQ, IWM, TLT, GLD, BTC-USD), 10 years daily data.
Standardized 5-day log returns form the 6-channel topographic signal.
$K$-means ($K=6$, trained on first temporal half) clusters returns into
microstates. QCI computed with window $W=63$ trading days.

Target: future 21-day realized volatility of SPY.

\subsection{Results}

\begin{table}[h]
\centering
\caption{QCI vs.\ future volatility (out-of-sample, $n = 1203$ days)}
\begin{tabular}{lcc}
\toprule
Metric & Value & Significance \\
\midrule
QCI vs.\ future vol & $r = -0.32$ & $p < 10^{-6}$ \\
Partial $r$ (controlling lagged RV) & $-0.22$ & $p < 10^{-8}$ \\
Robustness grid (80 configs) & 67/80 significant & 84\% \\
Permutation test (1000 shuffles) & $\chi^2 = 23.48$ & $> \max(\text{null}) = 12.88$ \\
Early warning ($\chi^2$ test) & $\chi^2 = 155$ & precision 54\%, recall 48\% \\
\bottomrule
\end{tabular}
\end{table}

\textbf{Interpretation}: When cross-asset dynamics deviate from QA
Fibonacci-shift predictions (low QCI), future volatility increases.
The partial correlation demonstrates this is not merely a volatility
proxy --- QCI carries independent structural information.

\subsection{Regime Structure}

The orbit distribution differs significantly between stress and calm regimes
($\chi^2 = 9.34$, $p = 0.009$, out-of-sample, $K=6$).
T-operator prediction accuracy itself varies by regime:
calm markets show 21\% accuracy (above the 16.7\% chance level),
stress markets show 12.3\% (below chance).
This means calm markets are more QA-coherent --- their dynamics more closely
follow the Fibonacci shift --- while stress markets are QA-incoherent.

\subsection{Baseline Comparison}

\begin{table}[h]
\centering
\caption{Baselines: correlation with future 21-day vol (OOS)}
\begin{tabular}{lc}
\toprule
Feature & $r$ with future vol \\
\midrule
Lagged realized vol (21d) & $+0.50$ \\
QCI ($K=6$, $W=63$) & $-0.32$ \\
Cross-asset correlation & $+0.25$ \\
PCA concentration & $+0.10$ \\
\midrule
QCI partial (controlling RV) & $-0.22$ \\
\bottomrule
\end{tabular}
\end{table}

We benchmark against standard volatility models on the same out-of-sample
period (601 test days):

\begin{table}[h]
\centering
\caption{OOS benchmark: future 21-day realized vol prediction ($n = 601$ test days)}
\begin{tabular}{llcc}
\toprule
Method & Features & OOS $R^2$ & OOS $r$ \\
\midrule
\multicolumn{4}{l}{\emph{Linear regression}} \\
 & HAR & $-0.048$ & 0.322 \\
 & HAR + QCI & $-0.017$ & 0.348 \\
\midrule
\multicolumn{4}{l}{\emph{GARCH(1,1)}} \\
 & conditional variance & $+0.030$ & 0.373 \\
\midrule
\multicolumn{4}{l}{\emph{Random Forest}} \\
 & HAR & $+0.183$ & 0.450 \\
 & HAR + QCI & $\mathbf{+0.202}$ & $\mathbf{0.474}$ \\
\midrule
\multicolumn{4}{l}{\emph{QCI standalone (linear)}} \\
 & QCI$_\text{gap}$ only & $-0.080$ & 0.440 \\
\bottomrule
\end{tabular}
\end{table}

The best model is Random Forest with HAR + QCI features ($R^2 = 0.20$,
$r = 0.47$), improving over RF with HAR alone ($R^2 = 0.18$,
$\Delta R^2 = +0.019$). QCI$_\text{gap}$ as a standalone linear predictor
achieves $r = 0.44$ --- competitive with RF HAR despite using no volatility
history. These results confirm that QCI carries genuine structural
information that complements classical volatility models.

\subsection{Multi-step T-prediction}

The T-operator can be applied $k$ times to predict the state $k$ steps ahead.
Testing $k \in \{1,2,3,5,8\}$:

\begin{table}[h]
\centering
\caption{Multi-step QCI: partial $r$ with future vol (controlling lagged RV, OOS)}
\begin{tabular}{ccc}
\toprule
$k$ & partial $r$ & Significance \\
\midrule
1 & $-0.215$ & $p < 10^{-8}$ \\
2 & $\mathbf{-0.260}$ & $p < 10^{-8}$ \\
3 & $-0.044$ & ns \\
5 & $+0.077$ & $p < 0.01$ \\
8 & $+0.134$ & $p < 10^{-3}$ \\
\bottomrule
\end{tabular}
\end{table}

The $k=2$ (double-step) lookahead is \emph{stronger} than $k=1$
(partial $r = -0.26$ vs $-0.22$), consistent with the algebraic
structure of the Fibonacci shift where $T^2$ has special properties.
The signal vanishes at $k=3$ and reverses sign at $k \geq 5$, indicating
a coherence horizon of approximately 2 transition steps.

\subsection{Nonlinear model comparison}

MLP neural networks ($\{64,32\}$ and $\{128,64,32\}$ architectures) do not
outperform Random Forest ($R^2 = 0.11$ vs $0.20$). The QCI signal is
adequately captured by shallow tree models, suggesting the information is
piecewise-linear rather than requiring deep nonlinear fitting.

\subsection{RV as Orbit Persistence}

When realized volatility is ingested as a 7th observer channel, the QCI
signal strengthens (raw $r = -0.44$ at $K=8$). This suggests that RV
persistence is a continuous projection of QA orbit persistence: the
T-operator naturally captures the state-transition autocorrelation that
underlies vol clustering. QA does not compete with classical observables ---
it provides the structural framework that explains them.

% ═══════════════════════════════════════════════════════════════════════════
\section{Cross-Domain Validation}
\label{sec:cross}
% ═══════════════════════════════════════════════════════════════════════════

\subsection{EEG Seizure Classification}

Using 23-channel EEG recordings from the CHB-MIT Scalp EEG Database
(10 patients), a topographic $k$-means observer classifies 1-second windows
into 4 microstates. QA orbit distributions differ significantly between
seizure and baseline: mean $\Delta R^2 = +0.21$, Fisher combined
$p = 2.9 \times 10^{-33}$, 10/10 patients significant.

The pattern: baseline EEG maps to singularity-dominated states (stable
resting topography), while seizure maps to cosmos-dominated states
(dynamic propagation). This mirrors the finance finding: instability
corresponds to departure from the singularity/dominant mode.

\subsection{Audio Signal Structure}

QA orbit transition rates differ by signal type: sine waves at 880~Hz
show elevated orbit return rates ($0.17$ vs $0.11$ chance), while white
noise matches the null distribution. Partial $r = +0.75$ ($p = 0.02$)
beyond lag-1 autocorrelation confirms QA structure is real, though the
effect is modest.

\subsection{Seismology: Earthquake Catalog}

Using the USGS earthquake catalog (152,237 M4+ events, 2016--2026),
daily features are computed: event count, mean magnitude, max magnitude,
and mean depth (4 channels). The topographic observer ($K=4$, mod-24)
applied to these features produces QCI that correlates with future
21-day seismic activity ($r = +0.23$, $p < 10^{-8}$, $n = 1679$ OOS days).

Orbit distributions differ significantly between active and quiet periods
($\chi^2 = 61.3$, $p < 10^{-6}$):
active periods show elevated satellite (60\%) and singularity (4\%)
rates, while quiet periods are cosmos-dominated (60\%).
This mirrors the finance pattern: instability corresponds to departure
from cosmos toward bounded/singularity orbits.

% ═══════════════════════════════════════════════════════════════════════════
\section{The Local/Global Observer Separation}
\label{sec:separation}
% ═══════════════════════════════════════════════════════════════════════════

Not all continuous observables should be naively stacked into the same
clustering space. We distinguish:

\begin{itemize}
  \item \textbf{Local state observables} (returns, RV, volume): describe the
    system's current coordinates. Safe to stack.
  \item \textbf{Global structure observables} (correlation, PCA eigenstructure,
    entropy): describe the geometry of the state space itself. Require a
    separate observer layer.
\end{itemize}

When cross-asset correlation was naively added as an 8th channel, the QCI
sign \emph{reversed} --- the correlation channel dominated clustering and
redefined what microstates represent. This is a \texttt{LOCAL\_GLOBAL\_LAYER\_COLLISION}
failure: a manifold descriptor was treated as a coordinate, changing the
semantic meaning of QCI without modeling it as a separate layer.

The correct architecture is a \emph{multi-layer observer}: local and global
observers produce separate QCI signals, and their discrepancy (the gap
between local coherence and global coherence) is where regime-change
information lives.

Empirical validation confirms this architecture. A pure global observer
using rolling correlation, PCA concentration, and dispersion features
produces $\text{QCI}_{\text{global}}$ with $r = +0.25$ with future volatility
(note the \emph{opposite sign} from local QCI). The gap signal
$\text{QCI}_{\text{gap}} = \text{QCI}_{\text{local}} - \text{QCI}_{\text{global}}$
achieves partial $r = -0.42$ beyond lagged RV at its best configuration ---
nearly double the local-only signal. Robustness: 16/16 grid configurations
significant for both global and gap signals (100\%).

The opposite signs carry the key insight: during stress, the coupling
structure \emph{tightens} (correlation increases $\to$ high global coherence)
while individual trajectories \emph{fragment} (chaotic movement $\to$ low
local coherence). The discrepancy between these layers --- structural
contraction with trajectory disorder --- is the universal stress signature.

% ═══════════════════════════════════════════════════════════════════════════
\section{The QA Observable Principle}
\label{sec:principle}
% ═══════════════════════════════════════════════════════════════════════════

We now state the theoretical principle underlying the empirical results.

\begin{definition}[QA Observer System]
A \emph{QA observer system} is a triple $(S_m, G, \pi)$ where:
\begin{itemize}
  \item $S_m = \{1,\ldots,m\}^2$ is the QA state space;
  \item $G = \{T\}$ is the generator set (Fibonacci shift, possibly extended
    with $\sigma, \mu, \lambda, \nu$);
  \item $\pi: S_m \to \mathcal{O}$ is an observer projection into a
    continuous observation space $\mathcal{O}$.
\end{itemize}
The \emph{reachability graph} $\Gamma_G$ has vertex set $S_m$ and directed
edges $(s, T(s))$ for each generator $T \in G$.
\end{definition}

\begin{theorem}[QA Observable Principle]
\label{thm:observable}
Let $(S_m, G, \pi)$ be a QA observer system. Then observed data
$o_0, o_1, \ldots$ are not primitive objects but arise as pushforwards of
path measures on $(S_m, \Gamma_G)$:
\[
  P_k(o \mid s_0) = \mu_k\bigl(\pi^{-1}(o)\bigr),
\]
where $\mu_k$ is a measure on states reachable from $s_0$ along paths of
length $k$ in $\Gamma_G$.

The \emph{QA Coherence Index} at time $t$ is the empirical estimate of
the probability that observed transitions conform to $\Gamma_G$:
\[
  \mathrm{QCI}_t = \frac{1}{W}\sum_{s=t-W+1}^{t}
    \mathbf{1}\bigl[(b_{s+1}, e_{s+1}) \in \Gamma_G(b_s, e_s)\bigr].
\]
\end{theorem}

\begin{proposition}[QCI as Manifold Deviation]
\label{prop:deviation}
Low QCI indicates that observed transitions frequently leave the admissible
edges of $\Gamma_G$. If the system's true dynamics are well-approximated
by QA generators, then low QCI signals departure from structural equilibrium.
Empirically, this manifests as:
\begin{enumerate}
  \item anticorrelation with future volatility (finance: $r = -0.32$);
  \item elevated cosmos-orbit rates (EEG seizure: $\Delta R^2 = +0.21$);
  \item reduced orbit return rates (audio: partial $r = +0.75$).
\end{enumerate}
\end{proposition}

% ═══════════════════════════════════════════════════════════════════════════
\section{Discussion}
\label{sec:discussion}
% ═══════════════════════════════════════════════════════════════════════════

\subsection{QA as a Measurement Framework}

The central claim of this paper is not that QA predicts specific outcomes
(return direction, seizure timing, signal identity). It is that
\emph{deviation from QA-native transition dynamics carries forward-looking
structural information about system stability}. QA provides a measurement
layer: a discrete algebraic framework into which continuous observables can
be ingested and structurally organized.

This positions QA alongside other structural frameworks:
\begin{itemize}
  \item In physics, symmetries describe system structure, not individual events.
  \item In control theory, state classification precedes control design.
  \item In QA, orbit coherence measures structural health before
    domain-specific prediction.
\end{itemize}

\subsection{Limitations}

\begin{enumerate}
  \item The microstate-to-QA mapping is designed, not learned. The mapping
    must ensure orbit diversity, which constrains the design space.
  \item The finance result is strongest at $K=5$--$6$ clusters and degrades
    at $K=4$. The sensitivity to cluster count suggests the signal is
    structured but not universal across all configurations.
  \item The multi-layer observer architecture (Section~\ref{sec:separation})
    is validated for finance but not yet tested cross-domain (EEG, audio).
  \item MLP neural networks underperform Random Forest on this dataset;
    transformer or attention-based architectures remain untested.
\end{enumerate}

\subsection{Future Work}

\begin{enumerate}
  \item Test multi-layer observer (local + global + fusion) cross-domain (EEG, audio).
  \item Apply to additional domains (climate, genomics, hyperspectral imaging).
  \item Investigate theoretical connections to phase-conjugate and pump-wave
    frameworks in nonlinear physics (the opposite-sign structure of local
    and global QCI bears resemblance to pump-conjugate dynamics).
\end{enumerate}

% ═══════════════════════════════════════════════════════════════════════════
\section{Conclusion}
% ═══════════════════════════════════════════════════════════════════════════

We have shown that the QA Fibonacci-shift operator, applied to topographically
clustered multi-channel signals, produces a coherence index with genuine
predictive power for future system instability. The result is hardened
across out-of-sample testing, robustness grids, permutation tests, and
partial correlation controls. Cross-domain validation in neuroscience and
audio processing demonstrates that the architecture is not domain-specific.

The deeper insight is foundational: continuous observables are pushforwards
of path measures on QA reachability graphs under observer projection.
QA does not approximate observables --- it defines the \emph{admissible
manifold} from which they are generated. The T-operator coherence measures
how closely a system's observed dynamics conform to the algebraic constraints
of this manifold. When conformity breaks down, the system is departing from
its own structural law --- and instability follows.

Unlike Koopman operator methods, which lift nonlinear dynamics to an
infinite-dimensional function space requiring truncation and approximation,
QA operates on a finite, algebraically closed state graph, enabling exact
reachability and obstruction analysis. This positions QA not as a feature
extractor but as a \emph{state-space lens}: a discrete algebraic framework
that constrains the problem \emph{before} learning happens.

\bigskip
\noindent\textbf{Data and code availability.}
All scripts (30--37) are frozen with SHA-256 hashes.
The QA certificate ecosystem (154 families) is available at
\texttt{github.com/1r0nw1ll/quantum-arithmetic-research}.

\end{document}
