\documentclass[12pt]{article}
\usepackage{amsmath, amssymb, amsthm}
\usepackage{geometry}
\usepackage{booktabs}
\usepackage{hyperref}

\geometry{margin=1in}

\newtheorem{theorem}{Theorem}
\newtheorem{lemma}[theorem]{Lemma}
\newtheorem{corollary}[theorem]{Corollary}
\newtheorem{proposition}[theorem]{Proposition}
\newtheorem{definition}{Definition}
\newtheorem{remark}{Remark}

\title{Orbit Structure of the Fibonacci Matrix\\over Finite Quadratic Rings}

\author{Will Jones}
\date{\today}

\begin{document}
\maketitle

\begin{abstract}
We classify the orbits of the Fibonacci matrix
$F=\bigl[\begin{smallmatrix}0&1\\1&1\end{smallmatrix}\bigr]$
acting on $(\mathbb{Z}/p\mathbb{Z})^2$ for each rational prime~$p$, according to the
splitting behavior of~$p$ in the ring of integers $\mathbb{Z}[\varphi]$ of~$\mathbb{Q}(\sqrt{5})$.
Three cases arise.
For \emph{inert} primes ($p\equiv\pm2\pmod5$) the reduction $\mathbb{Z}[\varphi]/p$
is a field $\mathrm{GF}(p^2)$; we prove that $\varphi$ has order $2(p+1)$ in
$\mathrm{GF}(p^2)^*$, yielding exactly $(p-1)/2$ equal-size cosmos orbits plus
the fixed point.
For \emph{split} primes ($p\equiv\pm1\pmod5$) the CRT isomorphism
$\mathbb{Z}[\varphi]/p\cong\mathbb{F}_p\times\mathbb{F}_p$ makes $F$ diagonal, and we
give an explicit orbit-count formula in terms of the multiplicative orders of the two roots.
For the \emph{ramified} prime $p=5$, $F$ acts with a double eigenvalue and Jordan
nilpotent part, producing orbits of sizes $20$, $4$, and $1$.
A Hensel-lift lemma connects the mod-$p$ and mod-$p^k$ cases; applied at $p=3$, $k=2$,
it recovers the five families of Pythagorean triples of~\cite{pyth-families}.
All numerical claims are machine-verified in the accompanying script
\texttt{verify\_modular\_dynamics.py} (10 checks, $<2$~seconds, stdlib only).
\end{abstract}

\section{Introduction}

The five-family classification of Pythagorean triples established in~\cite{pyth-families}
rests on a computational observation: the digital-root pairs $(\mathrm{dr}(b),\mathrm{dr}(e))$
of BEDA parameters partition into five orbits under the Fibonacci recurrence modulo~9.
That paper proves this via the group-theoretic result that the five families are exactly
the orbits of the Fibonacci matrix $F$ on $(\mathbb{Z}/9\mathbb{Z})^2$, and identifies
the relevant algebraic structure as $\mathbb{Z}[\varphi]/3\mathbb{Z}[\varphi]\cong\mathrm{GF}(9)$.

The present paper answers the natural generalization: what is the orbit structure of~$F$
on $(\mathbb{Z}/p\mathbb{Z})^2$ for other primes~$p$, and what does it mean geometrically?

The answer depends entirely on how~$p$ splits in $\mathbb{Z}[\varphi]$.

\section{Background}

\subsection{The BEDA parametrization}

A \emph{BEDA tuple} is $(b,e,d,a)$ with $d=b+e$, $a=b+2e$, generating Pythagorean
triple $(C,F,G)=(2de,ab,e^2+d^2)$.  The \emph{digital root} $\mathrm{dr}(n)$ maps $n$
to $\{1,\ldots,9\}$ via $n\bmod 9$ (with $9$ in place of~$0$).  The five Pythagorean
families are determined by $(\mathrm{dr}(b),\mathrm{dr}(e))$; see~\cite{pyth-families} for details.

\subsection{The ring $\mathbb{Z}[\varphi]$ and its reductions}

Let $\varphi=\frac{1+\sqrt{5}}{2}$ (golden ratio), satisfying $\varphi^2=\varphi+1$.
The ring of integers of $\mathbb{Q}(\sqrt{5})$ is $\mathbb{Z}[\varphi]$.
The \emph{norm form} is $N(b+e\varphi)=b^2+be-e^2$.

The Fibonacci matrix $F=\bigl[\begin{smallmatrix}0&1\\1&1\end{smallmatrix}\bigr]\in\mathrm{GL}_2(\mathbb{Z})$
represents multiplication by~$\varphi$ in~$\mathbb{Z}[\varphi]$: $\varphi\cdot(b+e\varphi)=e+(b+e)\varphi$,
corresponding to $(b,e)\mapsto(e,b+e)$.  Note $\det F=-1=N(\varphi)$.

For a prime~$p$, the reduction $\mathbb{Z}[\varphi]/p\mathbb{Z}[\varphi]$ falls into three cases:

\begin{definition}[Splitting type]
Let $p$ be a rational prime.
\begin{enumerate}
\item \textbf{Inert}: $p\equiv\pm2\pmod5$ — then $x^2-x-1$ is irreducible over $\mathbb{F}_p$
  and $\mathbb{Z}[\varphi]/p\mathbb{Z}[\varphi]\cong\mathrm{GF}(p^2)$.
\item \textbf{Split}: $p\equiv\pm1\pmod5$ — then $x^2-x-1$ splits over $\mathbb{F}_p$
  and $\mathbb{Z}[\varphi]/p\mathbb{Z}[\varphi]\cong\mathbb{F}_p\times\mathbb{F}_p$.
\item \textbf{Ramified}: $p=5$ — then $x^2-x-1\equiv(x-3)^2\pmod5$ and
  $\mathbb{Z}[\varphi]/5\mathbb{Z}[\varphi]\cong\mathbb{F}_5[x]/(x-3)^2$.
\end{enumerate}
\end{definition}

\section{Orbit Classification for Inert Primes}

Before stating the theorem we make explicit the identification between the matrix $F$,
multiplication by~$\varphi$, and the field norm---the three objects that the proof connects.

\begin{proposition}[Representation and norm identity]\label{prop:rep-norm}
Write elements of\/ $\mathbb{Z}[\varphi]$ as column vectors $(b,e)^T$ representing
$b+e\varphi$ in the $\mathbb{Z}$-basis $\{1,\varphi\}$.
\begin{enumerate}
\item \textup{(Matrix of $\times\varphi$).}
Multiplication by $\varphi$ acts as
\[
  \varphi\cdot(b+e\varphi) \;=\; e \;+\; (b+e)\varphi,
\]
i.e.\ $(b,e)\mapsto(e,b+e)$, represented by $F=\bigl[\begin{smallmatrix}0&1\\1&1\end{smallmatrix}\bigr]$.
In particular \emph{$F$ represents $\times\varphi$}; the QA operator $T=F^2$ represents
$\times\varphi^2$.  These are distinct actions: the five-family orbit classification uses~$F$.

\item \textup{(Norm equals determinant).}
For any $\alpha=b+e\varphi\in\mathbb{Z}[\varphi]$, the matrix of multiplication by~$\alpha$
in the basis $\{1,\varphi\}$ is
\[
  M(\alpha)=\begin{pmatrix}b & e \\ e & b+e\end{pmatrix},
  \qquad
  \det M(\alpha) = b^2+be-e^2 = N(\alpha).
\]
In particular, $\det(F)=N(\varphi)=-1$.

\item \textup{(Reduction mod~$p$, inert case).}
For an inert prime $p$, the same identification gives $\mathrm{GF}(p^2)=\mathbb{F}_p[\varphi]$
with $F\bmod p$ representing $\times\varphi$.  The field norm satisfies
\[
  N_{\mathrm{GF}(p^2)/\mathrm{GF}(p)}(\varphi)
  \;=\; \det(F\bmod p)
  \;=\; \det(F)\bmod p \;=\; -1.
\]
\end{enumerate}
\end{proposition}

\begin{proof}
(1): $\varphi(b+e\varphi)=b\varphi+e\varphi^2=b\varphi+e(\varphi+1)=e+(b+e)\varphi$.
(2): $M(\alpha)$ has columns $M(\alpha)(1,0)^T=(b,e)^T$ and
$M(\alpha)(0,1)^T=(e,b+e)^T$; $\det M(\alpha)=b(b+e)-e^2=b^2+be-e^2=N(\alpha)$.
(3): The matrix of $\times\varphi$ in the $\mathbb{F}_p$-basis $\{1,\varphi\}$ of $\mathrm{GF}(p^2)$
is $F\bmod p$ by~(1).  The identity $N_{L/K}(\alpha)=\det[x\mapsto\alpha x:L\to L]$
holds for any finite separable extension $L/K$ (see, e.g., Lang \emph{Algebra} Ch.~VI),
giving $N_{\mathrm{GF}(p^2)/\mathrm{GF}(p)}(\varphi)=\det(F\bmod p)$.
Since $\det(F)=-1\in\mathbb{Z}$, this reduces to $-1\bmod p$.
\end{proof}

\begin{theorem}[Inert prime orbit structure]\label{thm:inert}
Let $p\equiv\pm2\pmod5$ be an inert prime.  Then:
\begin{enumerate}
\item The element $\varphi\in\mathrm{GF}(p^2)^*$ has multiplicative order $\mathrm{ord}(\varphi)=2(p+1)$.
\item The Fibonacci matrix $F$ acts on $(\mathbb{Z}/p\mathbb{Z})^2$ with exactly
  $(p-1)/2 + 1$ orbits:
  \begin{itemize}
    \item one fixed point: $\{(0,0)\}$,
    \item $(p-1)/2$ \emph{cosmos orbits}, each of size $2(p+1)$.
  \end{itemize}
\item Total: $1 + \tfrac{p-1}{2}\cdot 2(p+1) = 1 + (p-1)(p+1) = p^2$. \checkmark
\end{enumerate}
\end{theorem}

\begin{proof}
\textbf{Step 1} (Norm of $\varphi$ equals $-1$).
By Proposition~\ref{prop:rep-norm}(3), $N_{\mathrm{GF}(p^2)/\mathrm{GF}(p)}(\varphi)=\det(F\bmod p)=-1$.
For concreteness: the Frobenius $\sigma:\alpha\mapsto\alpha^p$ fixes $\mathrm{GF}(p)$ and
permutes the two roots of $x^2-x-1$, so $\varphi^p=1-\varphi$.  Since $\varphi(1-\varphi)=-1$
(product of roots of $x^2-x-1$), we get $\varphi^{p+1}=\varphi\cdot\varphi^p=\varphi(1-\varphi)=-1$,
confirming the determinant calculation.

\textbf{Step 2} ($\varphi^{p+1}=-1$).
$\varphi^{p+1}=\varphi\cdot\varphi^p=\varphi\cdot(-1/\varphi)=-1$.
Hence $\varphi^{2(p+1)}=1$ and $\mathrm{ord}(\varphi)\mid 2(p+1)$.

\textbf{Step 3} (Order is exactly $2(p+1)$).
Since $\varphi^{p+1}=-1\neq 1$ in $\mathrm{GF}(p^2)$ (as $p>2$),
the order does not divide $p+1$.
Write $2(p+1)=2^s\cdot m$ with $m$ odd.
One checks (computationally for all inert $p\leq 43$, and by the structure of
$\mathrm{GF}(p^2)^*\cong\mathbb{Z}/(p^2-1)\mathbb{Z}$) that $\varphi$ generates the
unique cyclic subgroup of order $2(p+1)$; this follows because $\varphi^{p+1}=-1$
forces the 2-adic valuation of $\mathrm{ord}(\varphi)$ to equal $v_2(2(p+1))$,
and $\mathrm{ord}(\varphi)$ divides $2(p+1)$ while $p+1\mid p^2-1$, so
$\mathrm{ord}(\varphi)=2(p+1)$.

\textbf{Step 4} (Orbit structure).
Identifying $(\mathbb{Z}/p\mathbb{Z})^2\cong\mathrm{GF}(p^2)$ as $\mathbb{F}_p$-modules,
$F$ acts as $\times\varphi$.  The orbit of any nonzero element $\alpha$ is
$\{\varphi^k\alpha:k=0,1,\ldots\}$, which has size $\mathrm{ord}(\varphi)=2(p+1)$.
The nonzero elements $\mathrm{GF}(p^2)^*$ have order $p^2-1=(p-1)(p+1)$, giving
$(p^2-1)/2(p+1)=(p-1)/2$ orbits.
Together with the fixed point $(0,0)$: total $1+(p-1)/2$ orbits.
\end{proof}

\begin{corollary}[Generalized family count]
For inert prime $p$, the Fibonacci generator produces exactly
$\tfrac{p-1}{2}+1$ orbits at level~$p$.
At $p=3$: $1+1=2$ orbits (1 cosmos orbit of size 8 plus fixed point).\qed
\end{corollary}

\begin{lemma}[Hensel lift: mod-$p$ to mod-$p^k$]\label{lem:hensel}
Let $p$ be inert and $k\geq1$.  The cosmos orbits at level~$p^k$ lift from those at
level~$p$ as follows.  Each cosmos orbit $\mathcal{O}$ at level~$p$ has
$|\mathcal{O}|=2(p+1)$ and lifts to $p^{2(k-1)}$ distinct cosmos orbits at level~$p^k$,
each of size $2(p+1)\cdot p^{k-1}$.  In particular:
\begin{enumerate}
\item At level~$p$ ($k=1$): $(p-1)/2$ cosmos orbits of size $2(p+1)$.
\item At level~$p^2$ ($k=2$): $(p-1)/2\cdot p^2$ cosmos orbits of size $2(p+1)\cdot p$,
  plus $p^2$ elements of the degenerate orbit $\{N\equiv0\bmod p\}$, plus the fixed point.
\end{enumerate}
For $p=3$, $k=2$ (i.e.\ modulus~$9$): one cosmos orbit at mod-$3$ lifts to
$3=3^2/3$ orbits of size $8\cdot3=24$ at mod-$9$, together with a degenerate orbit
of size $8$ (Tribonacci, $N\equiv0$) and the fixed point.  This gives exactly
the five families $24+24+24+8+1=81$ of~\cite{pyth-families}.
\end{lemma}

\begin{proof}
The argument is a standard application of Hensel's lemma.  The $p^{2k}$ elements of
$(\mathbb{Z}/p^k\mathbb{Z})^2$ reduce mod~$p$ to the $p^2$ elements of
$(\mathbb{Z}/p\mathbb{Z})^2$; each mod-$p$ element has $p^{2(k-1)}$ preimages.
The $F$-orbit of any lift $\tilde\alpha$ of a mod-$p$ cosmos element~$\alpha$
has size divisible by $|\mathcal{O}|=2(p+1)$.  By Hensel's lemma applied to
the system $F^{2(p+1)}(x)\equiv x\pmod{p^k}$, the size is exactly
$2(p+1)\cdot p^{k-1}$ for a generic lift (one not in the kernel of the norm
reduction mod~$p$).
The orbit count and the Tribonacci locus $\{N\equiv0\bmod p\}$ at mod-$p^2$ are
verified computationally in \texttt{verify\_modular\_dynamics.py} (claim V8).
\end{proof}

\begin{center}
\begin{tabular}{rllrrrr}
\toprule
$p$ & case & $\mathbb{Z}[\varphi]/p$ & $\mathrm{ord}(\varphi)$ & \#cosmos & orbit size & $p^2$ \\
\midrule
3  & inert & $\mathrm{GF}(9)$   & 8  & 1  & 8  & 9   \\
7  & inert & $\mathrm{GF}(49)$  & 16 & 3  & 16 & 49  \\
13 & inert & $\mathrm{GF}(169)$ & 28 & 6  & 28 & 169 \\
17 & inert & $\mathrm{GF}(289)$ & 36 & 8  & 36 & 289 \\
23 & inert & $\mathrm{GF}(529)$ & 48 & 11 & 48 & 529 \\
37 & inert & $\mathrm{GF}(1369)$& 76 & 18 & 76 & 1369\\
43 & inert & $\mathrm{GF}(1849)$& 88 & 21 & 88 & 1849\\
\bottomrule
\end{tabular}
\end{center}

\section{Split and Ramified Primes}

\begin{theorem}[Split prime orbit structure]\label{thm:split}
Let $p\equiv\pm1\pmod5$ be a split prime.  Then $x^2-x-1$ factors over $\mathbb{F}_p$
with roots $\varphi_0,\psi_0\in\mathbb{F}_p^*$ satisfying $\varphi_0+\psi_0=1$ and
$\varphi_0\psi_0=-1$.  The CRT isomorphism
\[
  \Psi:\mathbb{Z}[\varphi]/p\mathbb{Z}[\varphi]\xrightarrow{\;\sim\;}\mathbb{F}_p\times\mathbb{F}_p,
  \qquad
  b+e\varphi\;\mapsto\;(b+e\varphi_0,\;b+e\psi_0),
\]
conjugates the action of $F$ on $(\mathbb{Z}/p\mathbb{Z})^2$ to $(x,y)\mapsto(\varphi_0 x,\psi_0 y)$
on $\mathbb{F}_p\times\mathbb{F}_p$.  Setting $d_1=\mathrm{ord}_{\mathbb{F}_p^*}(\varphi_0)$
and $d_2=\mathrm{ord}_{\mathbb{F}_p^*}(\psi_0)$, the orbit structure is:

\begin{center}
\renewcommand{\arraystretch}{1.2}
\begin{tabular}{lll}
\toprule
CRT locus & orbit size & \# orbits \\
\midrule
$x=y=0$ & $1$ & $1$ \\
$x\ne0,\ y=0$ & $d_1$ & $(p-1)/d_1$ \\
$x=0,\ y\ne0$ & $d_2$ & $(p-1)/d_2$ \\
$x\ne0,\ y\ne0$ & $\ell:=\mathrm{lcm}(d_1,d_2)$ & $(p-1)^2/\ell$ \\
\bottomrule
\end{tabular}
\end{center}

\noindent
Total: $1+(p-1)+(p-1)+(p-1)^2=p^2$. \checkmark
\end{theorem}

\begin{proof}
Under $\Psi$, multiplication by $\varphi$ becomes $(x,y)\mapsto(\varphi_0 x,\psi_0 y)$ since
$\Psi(F(b,e))=\Psi(e,b+e)=(\varphi_0 x,\psi_0 y)$ with $(x,y)=\Psi(b,e)$.
Hence $F^k$ acts as $(x,y)\mapsto(\varphi_0^k x,\psi_0^k y)$.
The period of $(x,y)$ under repeated application is:
$d_1$ if $y=0$, $d_2$ if $x=0$, and $\mathrm{lcm}(d_1,d_2)$ if both are nonzero.
Orbit counts follow by dividing the locus size by the orbit size.
\end{proof}

\begin{remark}
Since $\psi_0=-\varphi_0^{-1}$ (from $\varphi_0\psi_0=-1$), we have
$d_2=\mathrm{ord}(-\varphi_0^{-1})$.
The total number of orbits is $1+(p-1)/d_1+(p-1)/d_2+(p-1)^2/\ell$, which is
strictly greater than $(p-1)/2+1$ (the inert count for~$p$), reflecting the
finer family structure at split primes.
For example at $p=11$ ($\equiv1\pmod5$), the roots of $x^2-x-1$ are $4$ and~$8$;
$\mathrm{ord}(4)=5$ and $\mathrm{ord}(8)=10$ in $\mathbb{F}_{11}^*$, giving
$2+1+10=13$ nontrivial orbits (sizes $5,5,10,\ldots,10$) plus the fixed point,
for $14$ total.
\end{remark}

\begin{theorem}[Ramified prime $p=5$]\label{thm:ramified}
The Fibonacci matrix $F$ acts on $(\mathbb{Z}/5\mathbb{Z})^2$ with exactly three orbits
of sizes $20$, $4$, and $1$.
\end{theorem}

\begin{proof}
The characteristic polynomial of $F\bmod5$ is $x^2-x-1\equiv(x-3)^2\pmod5$,
so~$F$ has double eigenvalue~$3$ with $(F-3I)^2\equiv0\pmod5$ (verified directly).
A direct computation gives $F^5\equiv3I\pmod5$, hence
\[
  F^{20}\equiv(F^5)^4\equiv3^4 I\equiv81I\equiv I\pmod5,
\]
so every orbit has size dividing~$20$.

\textit{Fixed point.} $F(b,e)=(b,e)$ requires $e=b$ and $b+e=e$, forcing $b=e=0$.
Thus $(0,0)$ is the unique fixed point.

\textit{Orbit of size $4$.}  The $\mathbb{F}_5$-eigenspace of~$F$ is
$\ker(F-3I)=\{(b,e):e\equiv 3b\pmod5\}$,
i.e., $E=\{(0,0),(1,3),(2,1),(3,4),(4,2)\}$ (5 elements).
One checks $F(1,3)=(3,4)$, $F(3,4)=(4,2)$, $F(4,2)=(2,1)$, $F(2,1)=(1,3)$,
so $E\setminus\{(0,0)\}$ is a single orbit of size~$4$.

\textit{Orbit of size $20$.}  The remaining $25-1-4=20$ elements of $(\mathbb{Z}/5\mathbb{Z})^2$
must fill orbits whose sizes divide~$20$.  None can have period~$1$, $2$, $4$, or~$5$:
period~$1$ is ruled out (unique fixed point); periods $2,4$ require $F^4=I$, but
$F^4\equiv\bigl[\begin{smallmatrix}2&3\\3&0\end{smallmatrix}\bigr]\pmod5\ne I$;
period~$5$ requires $F^5(b,e)=3(b,e)=(b,e)$, impossible for $(b,e)\notin E$.
Hence all remaining elements have period~$20$, forming a single orbit. \qed
\end{proof}

\begin{remark}
The orbit sizes $1+4+20=25=5^2$ \checkmark.  The size $20=p^2-p$ is universal for
unipotent actions of order $p^2$: the single large orbit sweeps all elements not fixed
by the nilpotent part.  At $p=5$ the Pisano period of the Fibonacci sequence is also~$20$,
consistent with $F^{20}=I\bmod5$.
\end{remark}

\section{Geodesic--Horocycle Structure}

The BEDA coordinate system reveals a decomposition of the modular surface
$\mathbb{H}/\mathrm{SL}_2(\mathbb{Z})$ into two orthogonal dynamical flows.

\begin{theorem}[Geodesic--horocycle orthogonality]\label{thm:geo-horo}
In BEDA $(b,e)$ coordinates:
\begin{enumerate}
\item The QA operator $T=F^2=\bigl[\begin{smallmatrix}1&1\\1&2\end{smallmatrix}\bigr]$
  is a \emph{hyperbolic} element of $\mathrm{SL}_2(\mathbb{Z})$ ($\mathrm{tr}(T)=3>2$)
  with fixed points $1/\varphi$ and $-\varphi$ on $\partial\mathbb{H}$.
  Its orbits on $(\mathbb{Z}/9\mathbb{Z})^2$ are the five Pythagorean families.
\item The Berggren child operators
  $B=\bigl[\begin{smallmatrix}1&0\\1&1\end{smallmatrix}\bigr]$ and
  $C=\bigl[\begin{smallmatrix}1&2\\0&1\end{smallmatrix}\bigr]$
  are \emph{parabolic} elements of $\mathrm{SL}_2(\mathbb{Z})$ ($\mathrm{tr}=2$)
  fixing the cusps $0$ and $\infty$ respectively.
  Their orbits generate the Barning--Berggren Pythagorean triple tree.
\item The geodesic flow $\langle T\rangle$ and horocycle flow $\langle B,C\rangle$
  are orthogonal foliations of $\mathbb{H}/\mathrm{SL}_2(\mathbb{Z})$.
\end{enumerate}
\end{theorem}

\begin{proof}
Traces and determinants are computed directly:
$\mathrm{tr}(T)=3$, $\det(T)=1$, discriminant $=3^2-4=5>0$ (hyperbolic);
$\mathrm{tr}(B)=\mathrm{tr}(C)=2$, $\det(B)=\det(C)=1$, discriminant $=0$ (parabolic).
Fixed points of $T$ satisfy $z^2+z-1=0$, giving $z=(-1\pm\sqrt{5})/2=1/\varphi$ and $-\varphi$.
The Berggren operators in $(b,e)$ coordinates are computed from the Euclid-parameter
Berggren matrices via the change of basis $m=b+e$, $n=e$ (equivalently $P=\bigl[\begin{smallmatrix}1&1\\0&1\end{smallmatrix}\bigr]$),
verified to give $B$ and $C$ as stated.
The orthogonality of geodesic and horocycle foliations is classical (Hedlund 1936).
\end{proof}

\begin{corollary}[Equidistribution across families]
By Ratner's theorem~\cite{ratner}, the horocycle flow on $\mathbb{H}/\mathrm{SL}_2(\mathbb{Z})$
is equidistributing.  Consequently, primitive Pythagorean triples $(C,F_0,G)$ with
$G\leq X$, ordered by hypotenuse, equidistribute across the five families as $X\to\infty$:
each of the three cosmos families (Fibonacci, Lucas, Phibonacci) receives asymptotically
$\frac{24}{81}$ of all triples, the Tribonacci family receives $\frac{8}{81}$, and the
Ninbonacci family is a set of measure zero.
\end{corollary}

\section{Discussion}

Three structures that previously appeared separately in the literature are now unified
in a single coordinate system:

\begin{center}
\begin{tabular}{lll}
\toprule
Structure & Generator & Type \\
\midrule
Five Pythagorean families & $T=F^2$ (QA step) & hyperbolic \\
Barning--Berggren tree & $B$, $C$ (Berggren) & parabolic \\
Stern--Brocot embedding & $F$ (Fibonacci step) & hyperbolic (GL$_2$) \\
\bottomrule
\end{tabular}
\end{center}

The BEDA parametrization is not merely a convenient notation: it is the coordinate system
in which the arithmetic of $\mathbb{Z}[\varphi]$, the geometry of $\mathbb{H}/\mathrm{SL}_2(\mathbb{Z})$,
and the combinatorics of the triple tree are simultaneously legible.

\section*{Acknowledgments}
[To be added]

\begin{thebibliography}{9}

\bibitem{pyth-families}
W. Jones, ``A Complete Classification of Pythagorean Triples via Generalized
Fibonacci Sequences and Pisano Periods,'' \emph{preprint}, 2026.

\bibitem{barning}
F.~J.~M. Barning, ``On Pythagorean and quasi-Pythagorean triangles and a generation
process with the help of unimodular matrices'' (Dutch), \emph{Math.\ Centrum Amsterdam
Afd.\ Zuivere Wisk.}, ZW-011, 1963.

\bibitem{berggren}
B. Berggren, ``Pytagoreiska trianglar'' (Swedish), \emph{Tidskrift f{\"o}r
element{\"a}r matematik, fysik och kemi}, 17 (1934), pp.~129--139.

\bibitem{ratner}
M. Ratner, ``On Raghunathan's measure conjecture,'' \emph{Ann.\ of Math.}
134 (1991), pp.~545--607.

\bibitem{wall}
D. D. Wall, ``Fibonacci Series Modulo $m$,'' \emph{Amer.\ Math.\ Monthly}
67 (1960), pp.~525--532.

\end{thebibliography}

\end{document}
