\documentclass[12pt,a4paper]{article}

\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage{lmodern}
\usepackage{amsmath,amssymb,amsthm}
\usepackage{geometry}
\usepackage{booktabs}
\usepackage{enumitem}
\usepackage{hyperref}
\usepackage{listings}
\usepackage{xcolor}

\geometry{margin=1in}

\lstset{
  basicstyle=\ttfamily\small,
  backgroundcolor=\color{gray!10},
  frame=single,
  breaklines=true,
  columns=fullflexible,
}

\newtheorem{definition}{Definition}
\newtheorem{proposition}{Proposition}
\newtheorem{postulate}{Postulate}

\title{A Quantum Arithmetic Completion Layer\\for Field Structure Theory\\[6pt]
\large Machine-Checkable Verification of Loop-Based Structural Physics}

\author{Will Dale\thanks{Quantum Arithmetic Research Group.
Corresponding author.} \\
Don Briddell\thanks{Field Structure Institute, Mt. Airy, Maryland, USA.}}

\date{April 2026 --- Revision v2}

\begin{document}

\maketitle

% ============================================================================
% ABSTRACT
% ============================================================================

\begin{abstract}
Theoretical frameworks that make numerically testable structural claims
deserve verification mechanisms commensurate with the specificity of those
claims. We present \textbf{Quantum Arithmetic (QA)} as a \emph{completion
layer}: a framework that formalizes state spaces, generators, invariants, and
failure modes without introducing new physical assumptions, and applies them
to the verification of an external target theory. We use \textbf{Field
Structure Theory (FST)} and its applied formulation \textbf{Structural
Physics (SP)}, due to Briddell, as a worked case study.
We formalize five explicit claims from the FST primary text: (i) the
Sierpinski Triangle Fractal (STF) decomposition of the proton's 1836 loops
into integer clusters; (ii) the lambda-to-proton decay bookkeeping
$2187 - 243 - 81 - 27 = 1836$ as exact integer arithmetic; (iii) a calibration
postulate $P_1$ identifying one loop with one electron rest mass as the
sanctioned bridge between the discrete QA layer and the continuous observer
layer; (iv) a purely geometric three-cluster quark partition that does not
assert Standard-Model flavor identification; and (v) the six-loop chiral pair
structure of first-generation fermions. The certificate bundle is validated
deterministically by a single Python validator and passes with three
logged warnings: one calibration drift for the lambda
($\Delta = 1.87\;\mathrm{MeV}$, $0.168\%$ vs PDG) and two in-source typos
in Briddell's Chapter 2 flagged for corrigendum. The direct-read proton-mass
calibration $\Pi(1836) = 938.194\;\mathrm{MeV}$ differs from the
PDG~2024 proton mass by $0.078\;\mathrm{MeV}$ ($0.0083\%$). All validation
satisfies the QA Observer Projection Firewall (Theorem NT): the discrete
state layer is integer-only, and the continuous observer layer is crossed
exactly once per calibration row via the declared projection $\Pi$.
This is a revision (v2) of the manuscript originally submitted as
Frontiers in Physics manuscript~1850870.
\end{abstract}

% ============================================================================
% 1. INTRODUCTION
% ============================================================================

\section{Introduction: Reproducibility of Structural Claims}

\subsection{Verification as a first-class concern}

A recurring challenge in theoretical physics is that the specificity of a
claim and the specificity of its verification can come apart. A proposal may
make quantitative predictions --- mass ratios, decay accounting, symmetry
thresholds --- but state them in prose rather than in a form that permits
\emph{deterministic replay} by an independent party. When this happens,
legitimate disagreement about the physics can be swamped by disagreement
about what the theory actually claims, because the theory's claims live in
prose and diagrams rather than in typed, checkable artifacts.

This is not a failure specific to any one theoretical program. It is a
reproducibility problem that cuts across theoretical physics generally,
from exploratory alternatives to frontier mainstream work. The standard of
verification we advocate is: given the same inputs and declared invariants,
any validator should reach the same conclusion about whether a stated claim
is internally consistent, and any discrepancy against an external reference
should be \emph{logged} rather than silently absorbed.

\subsection{Three levels of engagement}

We distinguish three levels of engagement with a theoretical proposal:
\begin{enumerate}[nosep]
  \item \textbf{Acceptance}: believing the theory is physically true.
  \item \textbf{Verification}: confirming that stated claims are internally
        consistent and reproducible.
  \item \textbf{Falsification}: identifying specific, bounded conditions
        under which a claim fails.
\end{enumerate}
Quantum Arithmetic, as used here, operates at levels (2) and (3). It does
not address level (1) --- physical truth is outside its scope. Its purpose
is to make structural theories \emph{auditable}.

\subsection{Contribution of this paper}

This paper does not propose new physics. It demonstrates how an existing
structural theory --- Field Structure Theory (FST) and its applied
formulation Structural Physics (SP), due to Briddell~\cite{briddellbook2020,
briddell2026} --- can be:
\begin{itemize}[nosep]
  \item \emph{formalized} via explicit integer state spaces, generators,
        invariants, and obstruction types;
  \item \emph{validated deterministically} via certificates with
        recomputable integer witnesses;
  \item \emph{bridged to the continuous observer layer} via a single
        declared observer projection $\Pi$ that crosses the discrete/continuous
        boundary exactly once per computation, enforcing the QA Observer
        Projection Firewall (Theorem NT);
  \item \emph{audited for numeric drift} against external references (PDG
        particle masses) in the observer layer only, with drift logged as
        structured data rather than either silently absorbed or promoted to
        hard failure.
\end{itemize}
The relationship is explicit: Briddell's companion paper presents the
theory; this paper completes it computationally.

\subsection{Revision note (v2)}

This is a revision of the manuscript originally submitted 2026-04-08
as Frontiers in Physics manuscript~1850870. A pre-submission re-audit
identified that the v1 certificate bundle contained: (a) a symmetry
witness (a hexagonal side-excess vector $[2,2,2,0,0,0]$) that is not
present in Briddell's primary source; (b) a quark loop-count
assignment inverted from the STF decomposition given in Briddell's
Fig.~2.64; and (c) a loop-to-MeV ``homomorphism'' that compared a
subtraction-path-computed proton mass ($936.322\;\mathrm{MeV}$)
against the PDG reference rather than using Briddell's own
direct-read value ($938.196\;\mathrm{MeV}$), producing a worse drift
than the calibration actually supports. The v2 mapping reported here
removes the unsourced witness, restores the STF-consistent partition
$1836 = 2 \cdot 729 + 378$, and reframes the loop-mass bridge as an
explicit calibration postulate $P_1$ routed through the declared
observer projection $\Pi$. Two in-source typos in Briddell's
Chapter~2 identified during the re-audit are flagged in
Section~\ref{sec:corrigendum} for corrigendum.

% ============================================================================
% 2. OVERVIEW OF FST
% ============================================================================

\section{Overview of Field Structure Theory}

We provide a brief, non-competitive summary of FST/SP sufficient for the
formalization that follows. For the full treatment, we refer the reader to
Briddell's companion manuscript~\cite{briddell2026} and book-length
treatment~\cite{briddellbook2020}.

\subsection{Core elements}

FST posits a non-local, space-filling structural potential called the
\textbf{Plenum}, composed of continuous action loops. Particles form when
these loops organize into stable structures through entanglement, rotation,
and condensation. Two complementary loop types operate simultaneously:
\textbf{Rspin} (counter-clockwise, anti-matter) and \textbf{Aspin}
(clockwise, real-matter). Rspin loops remain deployed and dynamically
rotating; Aspin loops may condense to form interaction boundaries.

\textbf{Cloaking} occurs when condensed Aspin loopage fully encloses
deployed Rspin loopage such that no Rspin segment intersects the
interaction boundary. In this state, Rspin is structurally present but
unobservable. The minimal space-defining unit is the \textbf{Structor}:
a six-loop nucleus formed from three deployed Rspin loops enclosed by
three condensed Aspin loops.

Table~\ref{tab:states} summarizes the sector and deployment state space.
Condensation is a deployment-state change within a fixed sector: an Aspin
loop that condenses remains Aspin, and chirality is preserved while
observability and boundary-forming role change.

\begin{table}[h]
\centering
\caption{FST loop state space: sector and deployment are independent.}
\label{tab:states}
\begin{tabular}{@{}llccc@{}}
\toprule
\textbf{State} & \textbf{Sector} & \textbf{Deployment} & \textbf{Observable?} & \textbf{Structor Role} \\
\midrule
Rspin (deployed) & Anti-matter & Deployed & Yes & Inner core (3 loops) \\
Aspin (deployed) & Real-matter & Deployed & Yes & --- \\
Aspin (condensed) & Real-matter & Condensed & No (boundary) & Outer shell (3 loops) \\
\bottomrule
\end{tabular}
\end{table}

\subsection{STF hierarchy and first-generation particles}

The framework maps loop counts onto the \textbf{Sierpinski Triangle
Fractal (STF)}, whose first seven iterations carry loop counts
$\{3, 9, 27, 81, 243, 729, 2187\}$. Briddell identifies these with the
first generation of the Standard Model
particles~\cite[Chap.~15]{briddellbook2020}. Crucially, Briddell observes
that ``the proton and neutron are not iterations of the STF although
their loop counts are well within the seven
iterations''~\cite[line 861]{briddellbook2020}: the proton at 1836 loops
lies \emph{between} iteration~6 (729 loops) and iteration~7 (2187 loops)
and is obtained by removing smaller nested STF sub-triangles from the
iteration-7 configuration. This is the decomposition we formalize in
Certificates I and II.

\subsection{Claims selected for completion}

We restrict our verification to the following explicit, numerically
testable claims from Briddell's primary source:
\begin{enumerate}[nosep]
  \item The STF decomposition of the proton's 1836 loops into integer
        clusters: $1836 = 729 + 729 + 243 + 81 + 27 + 27$, interpreted
        geometrically as two iteration-6 sub-triangles of 729 loops
        each plus one top cluster of 378 loops.
  \item The lambda-to-proton decay bookkeeping
        $2187 - 243 - 81 - 27 = 1836$ as exact integer arithmetic
        over the STF basis.
  \item A unit-identification postulate, declared by Briddell's own
        loop-to-MeV table, that one loop (one ``Love'') corresponds to
        one electron rest mass. This is a calibration postulate, not
        a derivation.
  \item The three-cluster partition $\{729, 729, 378\}$ of the proton
        as a geometric structural claim, disambiguated from any
        Standard-Model flavor identification.
  \item The six-loop chiral-pair structure of first-generation
        fermions: three Rspin plus three Aspin loops per Structor-based
        particle~\cite[line 795]{briddellbook2020}.
\end{enumerate}
Only these claims are addressed. We make no judgment on the broader
physical program of FST/SP.

% ============================================================================
% 3. QA AS A COMPLETION FRAMEWORK
% ============================================================================

\section{Quantum Arithmetic as a Completion Framework}

\subsection{Philosophy: verification without reinterpretation}

Quantum Arithmetic treats numbers as \emph{geometric objects with intrinsic
structure} rather than abstract quantities. In the context of this paper,
QA serves as a \emph{verification layer} --- it encodes the concepts of a
target theory (here, FST) into a formal structure that can be validated
deterministically, without reinterpreting or replacing the theory's own
ontology.

The core organizing principle of the QA certificate framework is:
\begin{equation}
  \mathrm{Capability} = \mathrm{Reachability}(S, G, I)
\end{equation}
where $S$ is a state space, $G$ is a generator set (permitted operations),
and $I$ is an invariant set (constraints that must hold). A claim is
\emph{verified} if a witness path exists from an initial state to a target
state using only generators in $G$ while preserving all invariants in $I$.
A claim \emph{fails} if an explicit obstruction prevents such a path, and
the failure is classified by type.

\subsection{The QA discrete layer and the observer projection firewall}
\label{sec:theorem-nt}

A distinguishing feature of the QA framework is that it draws a hard
line between the \emph{discrete QA layer} --- where all state is
integer-typed and all operations are integer arithmetic --- and the
\emph{continuous observer layer} --- where real-valued measurements,
MeV values, PDG references, and any floating-point arithmetic live.
These two layers are bridged at exactly one sanctioned boundary per
computation, via a declared \textbf{observer projection} $\Pi$.

\begin{definition}[Observer Projection Firewall]
An observer projection is a function $\Pi: S \to \mathbb{R}^n$ that maps
an integer state in the QA layer to a real-valued output in the observer
layer. The \emph{observer projection firewall} (Theorem~NT in the QA
axiom block) asserts: $\Pi$ is the \emph{only} direction of information
flow between the two layers. Continuous values are never causal inputs
to the QA layer; they may only be outputs. Any attempt to feed a
float-typed value back into QA state, a QA generator, or a QA invariant
is a Type-2b axiom violation.
\end{definition}

This firewall is the reason the v2 mapping in this paper differs from
the v1 mapping: the v1 validator compared MeV values to integer loop
ratios inside the same decision logic without declaring which side of
the firewall the comparison lived on, blurring the QA/observer
boundary. The v2 validator declares $\Pi$ explicitly, applies it
exactly $k$ times per certificate (where $k$ is the number of rows in
the calibration table), and all MeV arithmetic happens strictly on the
observer side of that projection.

\subsection{Formal objects}

For any target theory, a QA module spine introduces five formal objects:

\begin{definition}[Module Spine]
A \textbf{QA module spine} is a tuple
$\langle \Omega, \Sigma, \mathcal{G}, \mathcal{I}, \mathcal{F} \rangle$
where:
\begin{itemize}[nosep]
  \item $\Omega$ is the \textbf{pattern space} (configuration space of the
        target theory, restricted to the discrete QA layer);
  \item $\Sigma$ is a \textbf{state packet} (typed integer-valued fields);
  \item $\mathcal{G}$ is a \textbf{generator set} (permitted integer
        transformations);
  \item $\mathcal{I}$ is an \textbf{invariant set} (integer-valued
        quantities preserved or bounded);
  \item $\mathcal{F}$ is an \textbf{obstruction taxonomy} (classified
        failure modes, with hard failures vs logged warnings distinguished).
\end{itemize}
\end{definition}

For FST, the v2 instantiation is shown in Table~\ref{tab:spine}. Note
that all fields are integer-typed or enum-typed; no float-valued state
appears in the spine.

\begin{table}[h]
\centering
\caption{QA module spine (v2) for Field Structure Theory. All state fields
are integer-typed.}
\label{tab:spine}
\begin{tabular}{@{}lll@{}}
\toprule
\textbf{Object} & \textbf{Name} & \textbf{Description} \\
\midrule
\multicolumn{3}{@{}l}{\emph{Pattern Space}} \\
& Plenum $\Omega$ & Integer STF-basis decompositions of first-generation particle loop counts \\
\midrule
\multicolumn{3}{@{}l}{\emph{Generators (all integer-valued)}} \\
& $\sigma$ & Add one loop quantum: $\ell \to \ell + 1$ \\
& $\lambda_3$ & STF scale step: $\ell \to 3\ell$ \\
& $\mu$ & Chirality swap: Rspin $\leftrightarrow$ Aspin (identity on $\ell$) \\
& $\kappa$ & Condensation toggle: deploy $\leftrightarrow$ condense (identity on $\ell$) \\
& $\chi$ & Cloaking operator: modify observability (identity on $\ell$) \\
\midrule
\multicolumn{3}{@{}l}{\emph{Invariants}} \\
& $\ell_{\mathrm{total}}$ & Total integer loop count \\
& $\varphi_{\mathrm{STF}}$ & Integer STF basis coefficients, $\varphi \cdot [3,9,27,81,243,729,2187] = \ell_{\mathrm{total}}$ \\
& sector balance & For fermions, $|\text{Rspin}| = |\text{Aspin}| = 3$ \\
\midrule
\multicolumn{3}{@{}l}{\emph{Observer Projection (single firewall crossing)}} \\
& $\Pi$ & $\Pi(\ell) = \ell \cdot m_e$ where $m_e = 0.510998950\;69\;\mathrm{MeV}/c^2$ \\
\midrule
\multicolumn{3}{@{}l}{\emph{Obstruction Types}} \\
& \texttt{NOT\_IN\_STF\_BASIS} & Declared decomposition does not sum under integer arithmetic \\
& \texttt{SECTOR\_IMBALANCE} & Fermion state violates $|\text{Rspin}| = |\text{Aspin}| = 3$ \\
& \texttt{CLOAKED\_FROM\_OBSERVER} & Boundary intersection blocked for declared observer sector \\
& \texttt{CALIBRATION\_DRIFT} & Observer-layer WARNING: $\Pi(\ell)$ vs PDG drift \\
& \texttt{SOURCE\_INTERNAL\_INCONSISTENCY} & Observer-layer WARNING: in-source typo flagged \\
\bottomrule
\end{tabular}
\end{table}

\subsection{Deterministic replay and auditability}

All QA certificates satisfy a \emph{deterministic replay} contract:
\begin{enumerate}[nosep]
  \item All data is serialized as canonical JSON (UTF-8, sorted keys,
        no whitespace).
  \item Hashes are computed as SHA-256 of the canonical representation.
  \item A manifest records the hash of every artifact required for replay,
        including the validator source itself.
  \item Re-execution of the validator on the same artifacts must produce
        identical output.
\end{enumerate}
This is enforced by draft-07 JSON schemas and a validator whose source code
is itself hashed in the manifest.

% ============================================================================
% 4. POSTULATE P1 AND THE OBSERVER PROJECTION
% ============================================================================

\section{Postulate $P_1$ and the Observer Projection $\Pi$}
\label{sec:postulate}

Briddell's own shell-to-loop table~\cite[lines 1580--1587]{briddellbook2020}
maps seven STF iterations to specific MeV values:
\begin{center}
\begin{tabular}{rrrr}
\toprule
STF iteration & loops & Briddell MeV & ratio (MeV/loop) \\
\midrule
2 &  9 &    4.599 & 0.51100 \\
3 & 27 &   13.797 & 0.51100 \\
4 & 81 &   41.391 & 0.51100 \\
5 & 243 & 124.173 & 0.51100 \\
6 & 729 & 372.519 & 0.51100 \\
proton  & 1836 &  938.196 & 0.51100 \\
7 & 2187 & 1115.683 & 0.51016 \\
\bottomrule
\end{tabular}
\end{center}
The ratio is constant at $0.51100\;\mathrm{MeV/loop}$ across all
first-generation particles up to the proton, which coincides to four
decimal places with the electron rest mass from PDG~2024,
$m_e = 0.510998950\;69\;\mathrm{MeV}/c^2$. The seventh iteration
(lambda) departs slightly from this ratio; we return to this in
Section~\ref{sec:cert3}.

We formalize this as an explicit calibration postulate.

\begin{postulate}[$P_1$: Loop-mass calibration]
One loop (one ``Love'') is calibrated to one electron rest mass:
$1\;\text{Love} := m_e = 0.510998950\;69\;\mathrm{MeV}/c^2$.
\end{postulate}

$P_1$ is declared, not derived. It is the sole bridge between the integer
loop counts in the QA layer and MeV values in the observer layer. A reader
who rejects $P_1$ can read the remaining certificates in this paper as
integer-only structural claims with no MeV content; the validator's
integer arithmetic does not depend on $P_1$ and passes all structural
checks without it.

With $P_1$, the observer projection is:
\begin{equation}
  \Pi(\ell) = \ell \cdot m_e \quad (\text{int} \to \mathbb{R}, \text{one-way})
\end{equation}
The validator applies $\Pi$ exactly once per row of the calibration
table and strictly once per certificate claim, and all subsequent
float arithmetic (drift computation, tolerance comparison) lives on
the observer side of the firewall. The validator reports the total
count of firewall crossings per validation run as a metric
(\texttt{theorem\_nt\_firewall\_crossings}) for audit transparency;
for the v2 cert bundle this count is exactly 7 (one per calibration
row).

% ============================================================================
% 5. CERTIFICATE I: STF PROTON DECOMPOSITION
% ============================================================================

\section{Certificate I: STF Proton Decomposition}
\label{sec:cert1}

\subsection{Structural claim}

Briddell's Fig.~2.64~\cite[p. 107]{briddellbook2020} presents the
proton's 1836 loops as a Sierpinski-triangle decomposition: two
bottom sub-triangles of 729 loops each, plus a top cluster of 378
loops, where $378 = 243 + 81 + 27 + 27$. The formal integer claim is:
\begin{equation}
  1836 = 729 + 729 + 243 + 81 + 27 + 27
\end{equation}

\subsection{Witness and verification}

The witness is the integer partition $\varphi = (729, 729, 243, 81, 27, 27)$.
The validator performs three integer-arithmetic checks:
\begin{enumerate}[nosep]
  \item All partition elements are positive integers.
  \item All partition elements belong to the STF basis
        $\{3, 9, 27, 81, 243, 729, 2187\}$.
  \item $\sum_i \varphi_i = 1836$.
\end{enumerate}
All three checks pass. No float arithmetic is performed.

\subsection{In-source typo flag}

Briddell's caption at MS line 1286 writes the formula as
``$729 + 729 + 243 + 27 + 27 = 1836$'', which is missing the
$81$ term. The actual sum of this formula is $1755$, not $1836$.
The \emph{geometry} shown in Fig.~2.64 is correct; the typo is
only in the in-caption arithmetic. The validator records this
under \texttt{SOURCE\_INTERNAL\_INCONSISTENCY}:

\begin{lstlisting}
{
  "move": "ENFORCE_STF_DECOMPOSITION",
  "fail_type": "SOURCE_INTERNAL_INCONSISTENCY",
  "severity": "warning",
  "invariant_diff": {
    "location": "Briddell 2026 MS line 1286",
    "actual_sum": 1755,
    "diagnosis": "Caption missing 81 term; correct formula is 729+729+243+81+27+27=1836"
  }
}
\end{lstlisting}

We flag this as a corrigendum item in Section~\ref{sec:corrigendum}.
It does not affect the structural decomposition, which the validator
reconstructs directly from the partition vector rather than from the
caption formula.

% ============================================================================
% 6. CERTIFICATE II: LAMBDA DECAY BOOKKEEPING
% ============================================================================

\section{Certificate II: Lambda Decay Bookkeeping}
\label{sec:cert2}

\subsection{Structural claim}

The lambda hyperon is STF iteration~7 at 2187 loops. The proton is
obtained from the lambda by removing three smaller nested STF
sub-triangles corresponding to iterations 5, 4, and 3 (243, 81, and
27 loops respectively):
\begin{equation}
  2187 - 243 - 81 - 27 = 1836
\end{equation}

\subsection{Witness and verification}

The witness is the pair (initial $= 2187$, subtract $= [243, 81, 27]$).
The validator performs three integer-arithmetic checks:
\begin{enumerate}[nosep]
  \item The initial count belongs to the STF basis.
  \item Every element of the subtract list belongs to the STF basis.
  \item $2187 - (243 + 81 + 27) = 1836$ as exact integer arithmetic.
\end{enumerate}
All three checks pass.

\subsection{In-source typo flag}

Briddell's text at MS line 1292 states that ``Missing are 378 loops,
which are the other decay products of the lambda.'' This is an
arithmetic slip: $2187 - 1836 = 351$, not $378$. The value $378$
appears elsewhere on the same page as the loop count of the
down-quark cluster in the proton's three-cluster decomposition, and
appears to have been accidentally carried over. The correct decay
loop count is $351 = 243 + 81 + 27$, which decomposes cleanly into
three STF sub-basis elements. The validator logs this under
\texttt{SOURCE\_INTERNAL\_INCONSISTENCY} and flags it as corrigendum
item~2 in Section~\ref{sec:corrigendum}.

% ============================================================================
% 7. CERTIFICATE III: LOOP-MASS CALIBRATION
% ============================================================================

\section{Certificate III: Loop-Mass Calibration Postulate}
\label{sec:cert3}

\subsection{Structural claim}

Under Postulate $P_1$, every first-generation particle in Briddell's
table has a predicted mass $\Pi(\ell) = \ell \cdot m_e$. The
validator applies $\Pi$ to the seven rows of the calibration table
and compares each against a declared PDG reference value (where
available), logging drift as structured data.

\subsection{Witness and verification}

Table~\ref{tab:calibration} shows the computed calibration. All
$\Pi$ values are computed by the validator and must match the
declared \texttt{pi\_mev\_expected} values to within $10^{-6}$
MeV; the validator fails hard on any larger mismatch.

\begin{table}[h]
\centering
\caption{Loop-mass calibration under $P_1$. Columns: integer loop
count $\ell$; projected mass $\Pi(\ell) = \ell \cdot m_e$ (declared
observer layer); Briddell's table value; PDG~2024 reference;
absolute drift against PDG; relative drift in percent.}
\label{tab:calibration}
\small
\begin{tabular}{@{}lrrrrrr@{}}
\toprule
Label     & $\ell$ & $\Pi(\ell)$ (MeV) & Briddell & PDG ref   & $\Delta$ (MeV) & $\Delta/\text{ref}$ (\%) \\
\midrule
STF iter 2 &    9 &    4.59899 &    4.599 & ---         & ---      & ---     \\
STF iter 3 &   27 &   13.79697 &   13.797 & ---         & ---      & ---     \\
STF iter 4 &   81 &   41.39092 &   41.391 & ---         & ---      & ---     \\
STF iter 5 &  243 &  124.17275 &  124.173 & ---         & ---      & ---     \\
STF iter 6 &  729 &  372.51824 &  372.519 & ---         & ---      & ---     \\
proton     & 1836 &  938.19407 &  938.196 & 938.27209   & 0.07802  & 0.00831 \\
lambda     & 2187 & 1117.55471 & 1115.683 & 1115.68300  & 1.87171  & 0.16776 \\
\bottomrule
\end{tabular}
\end{table}

\subsection{Interpretation}

The calibration holds to a fraction of $10^{-4}$ across the first six
STF iterations and at the proton, then drifts at the lambda by
approximately $0.17\%$. This is an observable, logged prediction:
$P_1$ is an approximation that degrades at the highest-mass
first-generation particle.

The validator's tolerance policy is:
\begin{itemize}[nosep]
  \item If $|\Delta|/\text{ref} \leq 0.01\%$: no action.
  \item If $0.01\% < |\Delta|/\text{ref} \leq 1\%$: log a
        \texttt{CALIBRATION\_DRIFT} warning.
  \item If $|\Delta|/\text{ref} > 1\%$: hard failure
        (\texttt{CALIBRATION\_DRIFT\_HARD}).
\end{itemize}
The proton drift ($0.00831\%$) is under the warning threshold; the
lambda drift ($0.16776\%$) fires a warning but does not exceed the
hard-fail band. The validator output for the v2 cert bundle is
\texttt{PASS\_WITH\_WARNINGS}.

The two thresholds are chosen as author-declared tolerances. They
are not derived from external ground truth. The contribution of this
certificate is not that the thresholds are correct, but that every
drift value is recomputed, logged, and publicly auditable as
structured data: an independent reader can set different thresholds
and re-run the validator to see which rows pass or fail under them.

\subsection{Comparison with v1}

The v1 mapping computed the proton mass via the subtraction path
$1115.683 - 124.173 - 41.391 - 13.797 = 936.322\;\mathrm{MeV}$ and
logged the drift against PDG as $1.95\;\mathrm{MeV}$. This path
selected the worse of the two numbers available in Briddell's text.
The v2 direct-read $\Pi(1836) = 938.194\;\mathrm{MeV}$ has drift
$0.078\;\mathrm{MeV}$, a $25\times$ improvement, and has a cleaner
epistemic status: it is the direct application of the declared
postulate rather than an indirect chain of source numbers that
propagates Briddell's own rounding.

% ============================================================================
% 8. CERTIFICATE IV: QUARK GEOMETRIC STRUCTURE
% ============================================================================

\section{Certificate IV: Quark Geometric Structure}
\label{sec:cert4}

\subsection{Structural claim}

The proton's 1836 loops decompose as an integer three-cluster
partition $\{729, 729, 378\}$. Briddell labels these clusters
``up quark (two)'' and ``down quark (one),''
respectively~\cite[lines 1287--1288]{briddellbook2020}, but these
labels refer to geometric positions in the Sierpinski triangle
decomposition (bottom-two and top-one), not to Standard Model
flavor assignments.

\subsection{Witness and verification}

The validator checks $729 + 729 + 378 = 1836$ as integer arithmetic.
Pass.

\subsection{Why we do not certify an SM flavor identification}
\label{sec:no-sm-flavor}

Under Postulate $P_1$, the Briddell clusters correspond to masses
$\Pi(729) = 372.5\;\mathrm{MeV}$ and $\Pi(378) = 193.2\;\mathrm{MeV}$.
These do not match the Standard Model current-quark masses
($m_u \approx 2.16\;\mathrm{MeV}$, $m_d \approx 4.67\;\mathrm{MeV}$
per PDG~2024) by three orders of magnitude, and they also do not
match Standard Model constituent-quark masses (both approximately
$336\;\mathrm{MeV}$ in non-relativistic quark models). The FST
clusters are geometric constructs of Briddell's decomposition, not
Standard-Model quark-flavor objects, and we do not certify any
identification between them.

\subsection{On the numerological ratio observation}

Briddell's Chapter~2~\cite[lines 1313--1326]{briddellbook2020}
records the observation that the integer ratio $378 / 729 = 0.5185$
is numerically close to a Standard Model $u/d$ mass ratio
computed from specific MeV values: ``$2.7 - 0.148 = 2.552\;\mathrm{MeV}$''
and ``$4.8625 + 0.0625 = 4.925\;\mathrm{MeV}$''. These
are described in the primary source as adjustments within PDG margin
of uncertainty selected to land on the $378/729$ ratio. Under the
PDG~2024 central values ($u/d \approx 0.4625$), the loop ratio and
the Standard Model current-quark mass ratio differ by approximately
$0.056$, or about $12\%$ of the ratio magnitude. We record the
observation for completeness but do not certify it as a verified
structural claim of FST, because the primary source itself describes
the back-walk procedure by which the numerical match was obtained.

\subsection{Comparison with v1}

The v1 cert bundle asserted an $u/d$ loop ratio certification with
$u = 378$, $d = 729$, and certified the numeric match between the
back-fit MeV ratio and the loop ratio as a PASS at tolerance $10^{-3}$.
This certification was tautological in two independent ways: the MeV
values were selected to target the loop ratio (making the match true
by construction), and the quark-to-cluster assignment in v1 was
inverted from Briddell's Fig.~2.64 caption (which places the two
bottom 729-loop clusters as Briddell's ``up'' and the 378-loop top
cluster as his ``down''). Under the v1 assignment $u = 378$, $d = 729$,
the proton decomposition $u + u + d$ sums to $1485$ loops, not 1836.
The v2 mapping restores the STF-geometric assignment ($1836 = 2 \cdot 729
+ 378$) and removes the flavor identification.

% ============================================================================
% 9. CERTIFICATE V: SIX-LOOP CHIRAL STRUCTURE
% ============================================================================

\section{Certificate V: Fermion Six-Loop Chiral Structure}
\label{sec:cert5}

\subsection{Structural claim}

Briddell states: ``A fermion mass is composed of six loops, three of
which are Rspin real-loop quarks and three are Aspin anti-loops
quarks''~\cite[line 795]{briddellbook2020}. We formalize this as
a type-level invariant on the FST state packet.

\subsection{Witness and verification}

Every first-generation fermion state must carry
$|\text{Rspin}| = |\text{Aspin}| = 3$ as a type-level constraint.
The validator performs one integer-equality check on the declared
sector cardinality. Pass.

\subsection{Sector mass accounting}

Briddell further states~\cite[line 861]{briddellbook2020}: ``Rspin and
Aspin loopage taken together gives a mass total of zero. The chiral
loops have positive and negative values to create zero mass and achieve
the needed symmetry.'' This is a qualitative structural statement
about sign conventions in FST's own mass accounting; it is not a
numerical claim and we record it as context but do not compute a
numerical certificate for it.

% ============================================================================
% 10. THEOREM NT COMPLIANCE SUMMARY
% ============================================================================

\section{Theorem NT Compliance}

The v2 cert bundle satisfies the QA Observer Projection Firewall
(Theorem NT) under the following protocol:
\begin{itemize}[nosep]
  \item \textbf{State fields are integer-typed.} The spine declares
        \texttt{loop\_count\_total\_loves} as \texttt{int},
        \texttt{stf\_basis\_decomposition} as \texttt{vector\_int},
        \texttt{sector\_cardinality} as \texttt{dict\_int}. No float
        state appears anywhere in the state packet.
  \item \textbf{Generators are integer-valued.} $\sigma$, $\lambda_3$,
        $\mu$, $\kappa$, $\chi$ all operate on integer state. No
        generator accepts or produces float values.
  \item \textbf{Certificates I, II, IV, V are integer-only.} All
        arithmetic performed by the validator on these certificates
        is over positive integers. No float cast occurs.
  \item \textbf{Certificate III has exactly one firewall crossing per
        row.} The validator applies $\Pi$ once per row of the
        calibration table. All subsequent arithmetic (drift
        computation, tolerance comparison) lives on the observer side
        of the crossing. No float value re-enters the integer layer.
  \item \textbf{The validator reports the firewall crossing count as a
        first-class metric}
        (\texttt{theorem\_nt\_firewall\_crossings}) so that any
        reader can audit how many times the boundary was crossed per
        validation run. For the v2 cert bundle this value is~7.
\end{itemize}
The QA axiom linter (\texttt{tools/qa\_axiom\_linter.py --all}) reports
zero errors across the repository, including on
\texttt{qa\_alphageometry\_ptolemy/qa\_fst/}. The ecosystem-level
meta-validator (\texttt{qa\_meta\_validator.py}) reports the FST
module as \texttt{PASS\_WITH\_WARNINGS} under its two-phase gate.

% ============================================================================
% 11. CORRIGENDUM ITEMS FOR BRIDDELL CHAPTER 2
% ============================================================================

\section{Corrigendum Items for the Primary Source}
\label{sec:corrigendum}

Our re-audit identified two in-source arithmetic slips in Briddell's
Chapter~2. Neither affects the underlying structural claims; both
are flagged here for correction in the next revision of the primary
manuscript and are logged by the v2 validator under
\texttt{SOURCE\_INTERNAL\_INCONSISTENCY}.

\paragraph{Corrigendum item 1: Fig.~2.64 caption formula (MS line 1286).}
The current text reads ``The Proton mass, having 1836 loops
$(729+729+243+27+27=1836)$.'' The arithmetic in parentheses is
missing the $81$ term; the actual sum of $729+729+243+27+27$ is
$1755$, not $1836$. The STF-correct formula is
$729+729+243+81+27+27 = 1836$. The \emph{geometry} in the figure
is correct as drawn; only the in-caption arithmetic has the
omission.

\paragraph{Corrigendum item 2: Lambda decay product count
(MS line 1292).}
The current text reads ``Missing are 378 loops, which are the other
decay products of the lambda.'' The correct count is $351$ loops,
not $378$. By direct subtraction, $2187 - 1836 = 351$, and $351$
decomposes cleanly in the STF basis as $243 + 81 + 27$ (iterations
5, 4, and 3). The value $378$ is the loop count of the down-quark
cluster described earlier on the same page (cf. Fig.~2.64 caption);
the two quantities appear to have been accidentally interchanged.

Both items are arithmetic bookkeeping slips rather than conceptual
errors. The underlying STF geometry, the proton decomposition, and
the lambda decay structure are unaffected.

% ============================================================================
% 12. REPRODUCIBILITY
% ============================================================================

\section{Reproducibility and Tooling}

\subsection{Artifact set}

The complete set of artifacts required for replay is published at a
version-controlled repository and consists of:

\begin{table}[h]
\centering
\caption{Published artifact set (v2).}
\label{tab:artifacts}
\begin{tabular}{@{}ll@{}}
\toprule
\textbf{Artifact} & \textbf{Role} \\
\midrule
\texttt{qa\_fst\_module\_spine.json} & Module spine (v2 declarative contract) \\
\texttt{qa\_fst\_cert\_bundle.json} & Worked certificates (five claims) \\
\texttt{qa\_fst\_submission\_packet\_spine.json} & Companion paper posture lock \\
\texttt{qa\_fst\_manifest.json} & SHA-256 manifest over all artifacts \\
\texttt{qa\_fst\_validate.py} & Deterministic validator + self-test \\
\texttt{schemas/*.schema.json} & Draft-07 JSON schemas \\
\bottomrule
\end{tabular}
\end{table}

The manifest covers nine artifacts, each hashed both as raw file
bytes (for file-integrity replay) and as canonical JSON (for
semantic-identity replay). The manifest itself records a composite
hash over all entries.

\subsection{One-command verification}

The validator can be run standalone:
\begin{lstlisting}
python qa_fst_validate.py           # run self-test suite
python qa_fst_validate.py --all     # full behavioral validation
python qa_fst_validate.py --check-manifest  # fast CI gate
\end{lstlisting}
The validator's self-test runs ten internal checks and exits with
status 0 on full pass or a non-zero code on any failure. At the
repository level, the FST module is invoked as a subprocess by
\texttt{qa\_meta\_validator.py}, and its JSON output is checked for
structural conformance by the ecosystem-level gate:
\begin{lstlisting}
python qa_meta_validator.py
\end{lstlisting}
Any third party with access to the repository can reproduce the
result.

% ============================================================================
% 13. COMPLETION LAYER POSTURE
% ============================================================================

\section{Why This Is a Completion Layer, Not a Competing Theory}

\subsection{What QA does not do}

It is important to state explicitly what the QA completion layer
does \emph{not} claim:
\begin{itemize}[nosep]
  \item It does not propose alternative physics. The Plenum, loops,
        STF hierarchy, Rspin/Aspin sectors, and cloaking mechanism
        are all Briddell's constructs, used here exactly as stated.
  \item It does not reinterpret FST concepts. The generators
        ($\sigma, \lambda_3, \mu, \kappa, \chi$) are named translations
        of operations described in the source text.
  \item It does not ``correct'' the theory. Where numeric drift or
        in-source typos exist, they are logged as warnings and
        flagged for corrigendum, not absorbed or silently adjusted.
  \item It does not certify Standard-Model flavor identifications
        that Briddell's own text does not support (see
        Section~\ref{sec:no-sm-flavor}).
\end{itemize}

\subsection{What QA adds}

The completion layer adds four capabilities that did not previously
exist for FST:
\begin{enumerate}[nosep]
  \item \textbf{Integer-typed state:} structural claims are encoded as
        integer partitions and cardinality constraints that a
        validator can check by exact arithmetic.
  \item \textbf{Observer projection firewall:} the bridge between
        discrete loop counts and continuous MeV values is a single
        declared function $\Pi$, crossed a known number of times per
        validation, so that every MeV value in the paper has an
        auditable provenance trace back to an integer state.
  \item \textbf{Replayability:} any third party can recompute every
        metric from the published artifacts and verify hash integrity.
  \item \textbf{Structured failure:} when something does not match
        --- whether an internal inconsistency, a calibration drift,
        or an in-source typo --- the discrepancy is classified by type,
        bounded by magnitude, and recorded for inspection. Hard
        failures, calibration warnings, and corrigendum flags are
        separable categories with separate severities.
\end{enumerate}

This posture converts ambiguity into data. Disagreement about the
theory becomes disagreement about specific numerical values, specific
thresholds, and specific source lines --- all of which can be pointed
at directly.

% ============================================================================
% 14. GENERALIZATION
% ============================================================================

\section{Generalization and Future Work}

The QA completion-layer approach is not specific to FST. Any
structural theory that makes numerically testable claims can be
formalized via the same pattern:
\begin{enumerate}[nosep]
  \item Identify the theory's state space, permitted integer
        operations, and conserved integer quantities.
  \item Encode these as a QA module spine.
  \item If the theory makes claims about continuous observables,
        declare an explicit observer projection $\Pi$ and a
        calibration postulate that bridges the two layers, with
        firewall crossings counted and reported.
  \item Select explicit, bounded claims and construct certificates
        with integer-recomputable witnesses.
  \item Classify failures by type and log drift against external
        references strictly on the observer side of the firewall.
\end{enumerate}

For FST specifically, natural extensions include:
\begin{itemize}[nosep]
  \item Second-generation and third-generation particle certificates
        beyond the lambda.
  \item Extended decay-path bookkeeping certificates with multi-step
        STF partitions.
  \item Formal cloaking certificates that test the observability
        obstruction (\texttt{CLOAKED\_FROM\_OBSERVER}) against
        specific sector/cloak-state configurations.
  \item A neutrino-scale calibration postulate paralleling $P_1$,
        using Briddell's ``Neutrino Sequence'' scaling (factor
        $\approx 2.2 \times 10^6$ between the electron-neutrino and
        the electron).
\end{itemize}

% ============================================================================
% 15. CONCLUSION
% ============================================================================

\section{Conclusion}

Structural theories of physics deserve rigorous verification
mechanisms commensurate with the specificity of their claims.
Quantum Arithmetic provides a neutral, formal completion layer
that encodes claims as typed integer certificates, declares the
discrete/continuous boundary via an explicit observer projection
$\Pi$, validates them deterministically, and classifies failures
as structured data.

Applied to Field Structure Theory, the v2 completion layer
demonstrates that:
\begin{itemize}[nosep]
  \item the proton STF decomposition is internally consistent:
        $729 + 729 + 243 + 81 + 27 + 27 = 1836$;
  \item the lambda decay bookkeeping is exact:
        $2187 - 243 - 81 - 27 = 1836$;
  \item the loop-mass calibration postulate $P_1$ (one loop equals
        one electron rest mass) holds at the $10^{-4}$ level across
        the first six STF iterations and at the proton (drift
        $0.0083\%$ vs PDG), and drifts by $\approx 0.17\%$ at the
        lambda;
  \item the three-cluster proton partition $\{729, 729, 378\}$ is
        structurally consistent but does \emph{not} support a
        Standard-Model flavor identification, because the numerical
        ratio match is explicitly back-fit within PDG tolerance in
        Briddell's own primary source;
  \item the first-generation fermion invariant
        $|\text{Rspin}| = |\text{Aspin}| = 3$ holds;
  \item two in-source arithmetic typos are flagged for corrigendum
        without affecting the underlying structural claims.
\end{itemize}

The completion layer neither endorses nor disputes FST's physical
claims. It makes them \emph{auditable}. We invite independent
verification, extension, and --- where warranted --- precise,
structured disagreement.

% ============================================================================
% REFERENCES
% ============================================================================

\begin{thebibliography}{9}

\bibitem{briddellbook2020}
D.~Briddell,
\emph{The Particle Hierarchy Paradigm: According to Field Structure Theory},
Field Structure Institute, Mt.~Airy, MD, 2020.
ISBN 978-1-7357041-0-4. (First edition 2020; revised draft 2026.)

\bibitem{briddell2026}
D.~Briddell,
``The Plenum-to-Particle Loop Sequence of Field Structure Theory (FST),''
manuscript submitted to \emph{Frontiers in Physics}, 2026;
manuscript ID 1850870.

\bibitem{pdg2024}
S.~Navas \emph{et al.} (Particle Data Group),
``Review of Particle Physics,''
\emph{Phys.\ Rev.\ D} \textbf{110}, 030001 (2024).

\bibitem{qaresearch2026}
W.~Dale,
``Quantum Arithmetic Research: Certificate Spine and Completion Layer,''
Quantum Arithmetic Research Group, 2026.

\end{thebibliography}

% ============================================================================
% APPENDIX
% ============================================================================

\appendix

\section{Validator Output (v2)}

The complete JSON output of \texttt{qa\_fst\_validate.py --all --json}
for the v2 cert bundle is reproduced below.

\begin{lstlisting}
{
  "result": "PASS_WITH_WARNINGS",
  "ok": true,
  "hashes": {
    "spine": "2504d691a35dc53e...",
    "cert_bundle": "bbadfd389981e723..."
  },
  "metrics": {
    "proton_partition": [729, 729, 243, 81, 27, 27],
    "proton_partition_sum": 1836,
    "proton_declared_total": 1836,
    "lambda_initial_loops": 2187,
    "lambda_subtract_sum": 351,
    "lambda_decay_final_recomputed": 1836,
    "fermion_sector_rspin": 3,
    "fermion_sector_aspin": 3,
    "quark_partition": [729, 729, 378],
    "quark_partition_sum": 1836,
    "theorem_nt_firewall_crossings": 7,
    "calibration_drift_count": 1,
    "calibration_table_recomputed": [
      {"label": "STF iter 2", "loops": 9,    "pi_mev_recomputed": 4.59899055621},
      {"label": "STF iter 3", "loops": 27,   "pi_mev_recomputed": 13.79697166863},
      {"label": "STF iter 4", "loops": 81,   "pi_mev_recomputed": 41.39091500589},
      {"label": "STF iter 5", "loops": 243,  "pi_mev_recomputed": 124.17274501767},
      {"label": "STF iter 6", "loops": 729,  "pi_mev_recomputed": 372.51823505301},
      {"label": "proton",     "loops": 1836, "pi_mev_recomputed": 938.19407346684,
       "pdg_ref_value": 938.27208816, "drift_mev": 0.07801469316, "drift_percent": 0.00831},
      {"label": "lambda",     "loops": 2187, "pi_mev_recomputed": 1117.55470515903,
       "pdg_ref_value": 1115.683, "drift_mev": 1.87170515903, "drift_percent": 0.16776}
    ]
  },
  "warnings": [
    {
      "move": "ENFORCE_STF_DECOMPOSITION",
      "fail_type": "SOURCE_INTERNAL_INCONSISTENCY",
      "invariant_diff": {
        "location": "Briddell 2026 MS line 1286",
        "actual_sum": 1755,
        "diagnosis": "Caption missing 81 term; correct formula 729+729+243+81+27+27=1836"
      }
    },
    {
      "move": "ENFORCE_STF_DECOMPOSITION",
      "fail_type": "SOURCE_INTERNAL_INCONSISTENCY",
      "invariant_diff": {
        "location": "Briddell 2026 MS line 1292",
        "actual_missing": 351,
        "diagnosis": "2187-1836=351=243+81+27; the '378' on line 1292 is the down-quark cluster loop count carried over in error"
      }
    },
    {
      "move": "APPLY_PI",
      "fail_type": "CALIBRATION_DRIFT",
      "invariant_diff": {
        "label": "lambda",
        "loops": 2187,
        "pi_mev": 1117.55470515903,
        "pdg_ref": 1115.683,
        "drift_mev": 1.87170515903,
        "drift_percent": 0.16776
      }
    }
  ]
}
\end{lstlisting}

\section{SHA-256 Manifest}

The manifest covers nine artifacts (three data files, five schemas,
the validator source). Each artifact carries both a file-bytes hash
and a canonical-JSON hash (except the raw Python source, which is
file-bytes only). A representative excerpt of the v2 manifest:

\begin{lstlisting}
module_spine:      2504d691a35dc53e...
cert_bundle:       bbadfd389981e723...
submission_packet: caa31ec5ad4887be...
validator_source:  4de1ef4ebdf071e5...
manifest_sha256:   dcac6e9594ea36b839be9a42d166d16e9c5bd919f9fd16627c36ce9bc6926f8d
\end{lstlisting}

Full hashes and all schema entries are recorded in
\texttt{qa\_fst\_manifest.json}.

\end{document}
